### 2023-05-22 outputs/paper/paper.qmd
Updated the preamble with personal information, data, and information relevant to my paper.

---
title: "Do More Polling Locations in Wards Increase Turnout? An Analysis of the 2022 Toronto Municipal Election"
author: "Inessa De Angelis"
thanks: "Code and data from this analysis are available at: https://github.com/InessaDeAngelis/Toronto_Elections."
date: "`r Sys.time()`"
date-format: "22 05 2023"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: ref.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("outputs/models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2023-05-23 scripts/00-simulate_data.R
Updated preamble

@@ -1,19 +1,19 @@
 #### Preamble ####
-# Purpose: Download the 2022 Toronto Municipal Election results data set 
+# Purpose: Simulate the 2022 Toronto Municipal Election ward locations and results data sets 
 # Author: Inessa De Angelis
-# Date: 22 May 2023 
+# Date: 25 May 2023 
 # Contact: inessa.deangelis@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: none
-# Data sets: https://open.toronto.ca/dataset/elections-voting-locations/ 
-# Note: Only download 2022 election data from the City of Toronto's website
-
+# Data sets: https://open.toronto.ca/dataset/elections-voting-locations/ & https://open.toronto.ca/dataset/election-results-official/ & https://open.toronto.ca/dataset/elections-voter-statistics/
+# Note: For each webpage, only download the 2022 election data 
 
 #### Workspace setup ####
 library(tidyverse)
 
 #### Simulate data ####
-# [...ADD CODE HERE...]
+
+## Assumptions
 
 
 

### 2023-05-23 scripts/01-download_data.R
Added the additional two data sets (voting locations and poll by poll results) into the script, including updating with the raw data files and relevant package information from OpenDataToronto.

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Downloads and saves the data from OpenDataToronto portal
 # Author: Inessa De Angelis
-# Date: 22 May 2023
+# Date: 25 May 2023
 # Contact: inessa.deangelis@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: none
@@ -15,15 +15,39 @@ library(dplyr)
 #### Download data and write municipal election polling location data ####
 
 # download data
-raw_location_data = download_data_from_opendatatoronto
-package_id = "31dac8b2-2e15-4945-abef-ce98d248bb8e"
+raw_voting_locations = download_data_from_opendatatoronto (
+package_id = "31dac8b2-2e15-4945-abef-ce98d248bb8e",
 resource_id = "31dac8b2-2e15-4945-abef-ce98d248bb8e"
+)
 
 # write data
 write.csv (
-  x = raw_location_data,
-  file = "inputs/data/raw_location_data.csv"
+  x = raw_voting_locations,
+  file = "inputs/data/raw_voting_locations.csv"
 )
 
+#### Download data and write municipal election voter statistics data ####
 
-         
+# download data
+raw_voter_statistics = download_data_from_opendatatoronto (
+  package_id = "7dc606ab-f042-4d90-99e2-9247cb5953d2",
+)
+
+# write data
+write.csv (
+  x = raw_voter_statistics,
+  file = "inputs/data/raw_voter_statistics.csv"
+)
+
+#### Download data and write municipal election poll by poll data ####
+
+# download data
+Raw_Poll_By_Poll_All_Offices = download_data_from_opendatatoronto (
+  package_id = "96d35404-44d9-49d8-95bb-fb1e5489240d",
+)
+
+# write data
+write.xlxs (
+  x = Raw_Poll_By_Poll_All_Offices,
+  file = "inputs/data/_Poll_By_Poll_All_Offices.xlsx"        
+)

### 2023-05-23 scripts/02-data_cleaning.R
Added janitor package

@@ -1,13 +1,14 @@
 #### Preamble ####
 # Purpose: Cleans the raw municipal election data by...
 # Author: Inessa De Angelis
-# Date: 22 May 2023
+# Date: 25 May 2023
 # Contact: inessa.deangelis@mail.utoronto.ca 
 # License: MIT
 # Pre-requisites: none
 
 #### Workspace setup ####
 library(tidyverse)
+library(janitor)
 
 #### Clean data ####
 raw_data <- read_csv("inputs/data/plane_data.csv")

### 2023-05-23 scripts/02-data_cleaning.R
Added pre-requisite

@@ -4,7 +4,7 @@
 # Date: 25 May 2023
 # Contact: inessa.deangelis@mail.utoronto.ca 
 # License: MIT
-# Pre-requisites: none
+# Pre-requisites: 01-download_data.R
 
 #### Workspace setup ####
 library(tidyverse)

### 2023-05-23 outputs/paper/paper.qmd
Installed and loaded a number of required packages

@@
 date: "`r Sys.time()`"
-date-format: "22 05 2023"
+date-format: "25 05 2023"
 abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."

### 2023-05-23 scripts/00-simulate_data.R
began writing data expectations and figuring out which columns on which of the raw data spreadsheets would be applicable.

@@ -11,6 +11,8 @@
 #### Workspace setup ####
 library(tidyverse)
 
+#### Data expectations ####
+
 #### Simulate data ####
 
 ## Assumptions

### 2023-05-23 scripts/00-simulate_data.R
Wrote data expectations, began writing simulations for the data with fictional wards, locations, and turnouts. Tests run so far have not been successful but further code needs to be added before a full simulation can be attampted.

@@ -12,10 +12,36 @@
 library(tidyverse)
 
 #### Data expectations ####
+# There are a similar or identical number of polling locations per ward
+# The polling locations in each ward in equally spaced out within the ward boundaries
+# columns: ward, building_name, building_address, %_voted
 
 #### Simulate data ####
 
 ## Assumptions
 
+# Fictional ward, voting locations, and voter turnout
+sim_ward = c("Ward1", "Ward2", "Ward3", "Ward44")
+sim_locations = c(55, 52, 57, 56)
+sim_turnout = c("22%", "27%", "32%", "38%")
 
+## Create summary statistics of number of polling locations per ward and 
+## compare it with the voter turnout by ward
 
+# summarize polling locations by ward
+polling_locations_per_ward = 
+  simulated_data |>
+  group_by(ward) |>
+  count() |>
+  rename("num_locations" = "n")
+
+# add voter turnout column to the data
+locations_per_ward['turnout'] = sim_turnout
+
+# compute bps per ward
+locations_per_ward =
+  locations_per_ward |>
+  mutate(
+    bps_of_ward =
+      num_locations / population * 10000
+  )

### 2023-05-23 scripts/01-download_data.R
Removed the poll by poll data set as the document is in the incorrect format.

@@ -34,20 +34,7 @@ raw_voter_statistics = download_data_from_opendatatoronto (
 )
 
 # write data
-write.csv (
+write.csv(
   x = raw_voter_statistics,
   file = "inputs/data/raw_voter_statistics.csv"
-)
-
-#### Download data and write municipal election poll by poll data ####
-
-# download data
-Raw_Poll_By_Poll_All_Offices = download_data_from_opendatatoronto (
-  package_id = "96d35404-44d9-49d8-95bb-fb1e5489240d",
-)
-
-# write data
-write.xlxs (
-  x = Raw_Poll_By_Poll_All_Offices,
-  file = "inputs/data/_Poll_By_Poll_All_Offices.xlsx"        
 )

### 2023-05-23 scripts/04-model.R
Deleting this script as it is not needed for this paper

@@ -1,35 +1,0 @@
-#### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Inessa De Angelis
-# Date: 22 May 2023
-# Contact: inessa.deangelis@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: none
-
-#### Workspace setup ####
-library(tidyverse)
-library(rstanarm)
-
-#### Read data ####
-analysis_data <- read_csv("outputs/data/analysis_data.csv")
-
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
-
-
-#### Save model ####
-saveRDS(
-  first_model,
-  file = "outputs/models/first_model.rds"
-)
-
-

### 2023-05-24 scripts/01-download_data.R
Removed and re-wrote the code to download and write the voter statistics data as there was an error with finding the file. Same issue appears in the code.

@@ -26,15 +26,15 @@ write.csv (
   file = "inputs/data/raw_voting_locations.csv"
 )
 
-#### Download data and write municipal election voter statistics data ####
+#### Download data and write voter statistics data ####
 
 # download data
 raw_voter_statistics = download_data_from_opendatatoronto (
-  package_id = "7dc606ab-f042-4d90-99e2-9247cb5953d2",
+  package_id = "7dc606ab-f042-4d90-99e2-9247cb5953d2"
 )
 
 # write data
-write.csv(
+write.csv (
   x = raw_voter_statistics,
   file = "inputs/data/raw_voter_statistics.csv"
 )

### 2023-05-24 scripts/01-download_data.R
added quotations around "raw_voter_statistics" which got rid of the error message when running the script.

@@ -35,6 +35,6 @@ raw_voter_statistics = download_data_from_opendatatoronto (
 
 # write data
 write.csv (
-  x = raw_voter_statistics,
+  x = "raw_voter_statistics",
   file = "inputs/data/raw_voter_statistics.csv"
 )

### 2023-05-24 scripts/00-simulate_data.R
Removed unused dataset URL, updated spreadsheet column names, and updated fictional data to simulate.

@@ -5,7 +5,7 @@
 # Contact: inessa.deangelis@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: none
-# Data sets: https://open.toronto.ca/dataset/elections-voting-locations/ & https://open.toronto.ca/dataset/election-results-official/ & https://open.toronto.ca/dataset/elections-voter-statistics/
+# Data sets: https://open.toronto.ca/dataset/elections-voting-locations/ & https://open.toronto.ca/dataset/elections-voter-statistics/
 # Note: For each webpage, only download the 2022 election data 
 
 #### Workspace setup ####
@@ -14,23 +14,23 @@ library(tidyverse)
 #### Data expectations ####
 # There are a similar or identical number of polling locations per ward
 # The polling locations in each ward in equally spaced out within the ward boundaries
-# columns: ward, building_name, building_address, %_voted
+# columns: point_name, point_short_code, voter_count
 
 #### Simulate data ####
 
 ## Assumptions
 
-# Fictional ward, voting locations, and voter turnout
-sim_ward = c("Ward1", "Ward2", "Ward3", "Ward44")
-sim_locations = c(55, 52, 57, 56)
-sim_turnout = c("22%", "27%", "32%", "38%")
+# Fictional point name, point short code, and voter count
+sim_point_name = c("Catholic_Church", "High_School", "Apartments", "Nursing_Home")
+sim_point_short_code = c(042, 025, 033, 060)
+sim_voter_count = c("1000", "1700", "500", "2100")
 
 ## Create summary statistics of number of polling locations per ward and 
 ## compare it with the voter turnout by ward
 
 # summarize polling locations by ward
 polling_locations_per_ward = 
-  simulated_data |>
+  "simulated_data" |>
   group_by(ward) |>
   count() |>
   rename("num_locations" = "n")

### 2023-05-24 scripts/00-simulate_data.R
Added "id" column to script.

@@ -14,13 +14,14 @@ library(tidyverse)
 #### Data expectations ####
 # There are a similar or identical number of polling locations per ward
 # The polling locations in each ward in equally spaced out within the ward boundaries
-# columns: point_name, point_short_code, voter_count
+# columns: _id, point_name, point_short_code, voter_count
 
 #### Simulate data ####
 
 ## Assumptions
 
 # Fictional point name, point short code, and voter count
+sim_id = c("1", "2", "3", "4",)
 sim_point_name = c("Catholic_Church", "High_School", "Apartments", "Nursing_Home")
 sim_point_short_code = c(042, 025, 033, 060)
 sim_voter_count = c("1000", "1700", "500", "2100")
@@ -30,7 +31,7 @@ sim_voter_count = c("1000", "1700", "500", "2100")
 
 # summarize polling locations by ward
 polling_locations_per_ward = 
-  "simulated_data" |>
+  simulated_data |>
   group_by(ward) |>
   count() |>
   rename("num_locations" = "n")

### 2023-05-24 scripts/00-simulate_data.R
Removed extra comma at the end of sim_id which was causing problems.

@@ -21,7 +21,7 @@ library(tidyverse)
 ## Assumptions
 
 # Fictional point name, point short code, and voter count
-sim_id = c("1", "2", "3", "4",)
+sim_id = c("1", "2", "3", "4")
 sim_point_name = c("Catholic_Church", "High_School", "Apartments", "Nursing_Home")
 sim_point_short_code = c(042, 025, 033, 060)
 sim_voter_count = c("1000", "1700", "500", "2100")

### 2023-05-24 scripts/00-simulate_data.R
Removed the voter statistics data set from the script

@@ -5,8 +5,8 @@
 # Contact: inessa.deangelis@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: none
-# Data sets: https://open.toronto.ca/dataset/elections-voting-locations/ & https://open.toronto.ca/dataset/elections-voter-statistics/
-# Note: For each webpage, only download the 2022 election data 
+# Data sets: https://open.toronto.ca/dataset/elections-voting-locations/ 
+# Note: Only download the 2022 election data from the webpage
 
 #### Workspace setup ####
 library(tidyverse)

### 2023-05-24 scripts/04-helper_functions.R
wrote preamble and added a function to download data from OpenDataToronto's website - hopefully fixing the challenges I have faced with my "download" script.

@@ -1,37 +1,19 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Author: Inessa De Angelis
+# Date: 25 May 2023
+# Contact: inessa.deangelis@mail.utoronto.ca 
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: none
 
 #### Workspace setup ####
-library(tidyverse)
-library(rstanarm)
 
-#### Read data ####
-analysis_data <- read_csv("outputs/data/analysis_data.csv")
+library(opendatatoronto)
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
-
-
-#### Save model ####
-saveRDS(
-  first_model,
-  file = "outputs/models/first_model.rds"
-)
-
-
+## Download data from opendatatoronto
+download_data_from_opendatatoronto = function(package_id, resource_id) {
+  
+  raw_data = 
+    list_package_resources(package_id) |> # get resources for the package_id
+    filter (id == resource_id) |> # filter by resource_id
+    get_resource() #download data
+}

### 2023-05-24 scripts/01-download_data.R
added helper function script to download data

@@ -11,6 +11,8 @@
 library(opendatatoronto)
 library(tidyverse)
 library(dplyr)
+library(readr)
+source("scripts/04-helper_functions.R")
 
 #### Download data and write municipal election polling location data ####
 
@@ -24,17 +26,4 @@ resource_id = "31dac8b2-2e15-4945-abef-ce98d248bb8e"
 write.csv (
   x = raw_voting_locations,
   file = "inputs/data/raw_voting_locations.csv"
-)
-
-#### Download data and write voter statistics data ####
-
-# download data
-raw_voter_statistics = download_data_from_opendatatoronto (
-  package_id = "7dc606ab-f042-4d90-99e2-9247cb5953d2"
-)
-
-# write data
-write.csv (
-  x = "raw_voter_statistics",
-  file = "inputs/data/raw_voter_statistics.csv"
 )

### 2023-05-24 scripts/00-simulate_data.R
updated column names

@@ -36,8 +36,8 @@ polling_locations_per_ward =
   count() |>
   rename("num_locations" = "n")
 
-# add voter turnout column to the data
-locations_per_ward['turnout'] = sim_turnout
+# add point name column to the data
+locations_per_ward['turnout'] = sim_point_turnout
 
 # compute bps per ward
 locations_per_ward =
@@ -46,3 +46,4 @@ locations_per_ward =
     bps_of_ward =
       num_locations / population * 10000
   )
+

### 2023-05-25 outputs/paper/paper.qmd
Made changes to the paper setup and removed example text.

@@
 thanks: "Code and data from this analysis are available at: https://github.com/InessaDeAngelis/Toronto_Elections."
-date: "`r Sys.time()`"
-date-format: "25 05 2023"
+date: "today"
+date-format: "long" 
 abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
@@
 ```{r}
-#| include: false
-#| warning: false
-#| message: false
 
-library(tidyverse)
-library(palmerpenguins)
 ```
-
-# Introduction
-
-You can and should cross-reference sections and sub-sections.
-
-The remainder of this paper is structured as follows. @sec-data....
-
-# Data {#sec-data}
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
-
-```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
-Talk way more about it.
-
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("outputs/models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
-
-# Discussion
-
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
-
-\newpage
-
-\appendix
-
-# Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
-
-\newpage
-
-# References

### 2023-05-25 scripts/01-download_data.R
re-added previously removed code for voter statistics

@@ -6,7 +6,6 @@
 # License: MIT
 # Pre-requisites: none
 
-
 #### Workspace setup ####
 library(opendatatoronto)
 library(tidyverse)
@@ -18,12 +17,25 @@ source("scripts/04-helper_functions.R")
 
 # download data
 raw_voting_locations = download_data_from_opendatatoronto (
-package_id = "31dac8b2-2e15-4945-abef-ce98d248bb8e",
-resource_id = "31dac8b2-2e15-4945-abef-ce98d248bb8e"
+  package_id = "31dac8b2-2e15-4945-abef-ce98d248bb8e",
+  resource_id = "31dac8b2-2e15-4945-abef-ce98d248bb8e"
 )
 
 # write data
 write.csv (
   x = raw_voting_locations,
   file = "inputs/data/raw_voting_locations.csv"
+)
+
+#### Download data and write municipal election voter statistics data ####
+
+# download data
+raw_voter_statistics = download_data_from_opendatatoronto (
+  package_id = "7dc606ab-f042-4d90-99e2-9247cb5953d2",
+)
+
+# write data
+write.csv (
+  x = raw_voter_statistics,
+  file = "inputs/data/raw_voter_statistics.csv"
 )

### 2023-05-25 scripts/04-helper_functions.R
added in resource_id

@@ -10,10 +10,9 @@
 library(opendatatoronto)
 
 ## Download data from opendatatoronto
-download_data_from_opendatatoronto = function(package_id, resource_id) {
+download_data_from_opendatatoronto = function(package_id) {
   
   raw_data = 
     list_package_resources(package_id) |> # get resources for the package_id
-    filter (id == resource_id) |> # filter by resource_id
     get_resource() #download data
 }

### 2023-05-25 scripts/00-simulate_data.R
updated relevant information and removed example code.

@@ -5,15 +5,17 @@
 # Contact: inessa.deangelis@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: none
-# Data sets: https://open.toronto.ca/dataset/elections-voting-locations/ 
-# Note: Only download the 2022 election data from the webpage
+# Data sets:
+## - https://open.toronto.ca/dataset/elections-voting-locations/ 
+## - https://open.toronto.ca/dataset/elections-voter-statistics/ 
+# Note: Only download the 2022 election data from the webpages
 
 #### Workspace setup ####
+# install.packages("tidyverse")
 library(tidyverse)
 
 #### Data expectations ####
 # There are a similar or identical number of polling locations per ward
-# The polling locations in each ward in equally spaced out within the ward boundaries
 # columns: _id, point_name, point_short_code, voter_count
 
 #### Simulate data ####
@@ -22,28 +24,14 @@ library(tidyverse)
 
 # Fictional point name, point short code, and voter count
 sim_id = c("1", "2", "3", "4")
-sim_point_name = c("Catholic_Church", "High_School", "Apartments", "Nursing_Home")
+sim_point_name = c("Church", "High_School", "Apartments", "Nursing_Home")
 sim_point_short_code = c(042, 025, 033, 060)
 sim_voter_count = c("1000", "1700", "500", "2100")
 
-## Create summary statistics of number of polling locations per ward and 
-## compare it with the voter turnout by ward
+## Create summary statistics of number of polling locations per ward 
 
 # summarize polling locations by ward
-polling_locations_per_ward = 
+locations_per_ward = 
   simulated_data |>
   group_by(ward) |>
   count() |>
-  rename("num_locations" = "n")
-
-# add point name column to the data
-locations_per_ward['turnout'] = sim_point_turnout
-
-# compute bps per ward
-locations_per_ward =
-  locations_per_ward |>
-  mutate(
-    bps_of_ward =
-      num_locations / population * 10000
-  )
-

### 2023-05-25 scripts/04-helper_functions.R
updated with resource_id

@@ -10,7 +10,7 @@
 library(opendatatoronto)
 
 ## Download data from opendatatoronto
-download_data_from_opendatatoronto = function(package_id) {
+download_data_from_opendatatoronto = function(package_id, resource_id) {
   
   raw_data = 
     list_package_resources(package_id) |> # get resources for the package_id

### 2023-05-25 scripts/00-simulate_data.R
Removed previous assumptions, code, and updated columns

@@ -15,23 +15,9 @@
 library(tidyverse)
 
 #### Data expectations ####
-# There are a similar or identical number of polling locations per ward
-# columns: _id, point_name, point_short_code, voter_count
+# There are a similar or identical number of subdivisions per ward
+# columns: ward, sub
 
 #### Simulate data ####
 
-## Assumptions
 
-# Fictional point name, point short code, and voter count
-sim_id = c("1", "2", "3", "4")
-sim_point_name = c("Church", "High_School", "Apartments", "Nursing_Home")
-sim_point_short_code = c(042, 025, 033, 060)
-sim_voter_count = c("1000", "1700", "500", "2100")
-
-## Create summary statistics of number of polling locations per ward 
-
-# summarize polling locations by ward
-locations_per_ward = 
-  simulated_data |>
-  group_by(ward) |>
-  count() |>

### 2023-05-26 scripts/01-download_data.R
The read_csv code was producing errors, so using the following blog post: https://community.rstudio.com/t/read-csv-doesnt-work-in-r-studio/116084 I updated the code to use the "readr" package to read in the raw code already in the "inputs" folder (which had been downloaded from OpenDataToronto's website).

@@ -15,27 +15,8 @@ source("scripts/04-helper_functions.R")
 
 #### Download data and write municipal election polling location data ####
 
-# download data
-raw_voting_locations = download_data_from_opendatatoronto (
-  package_id = "31dac8b2-2e15-4945-abef-ce98d248bb8e",
-  resource_id = "31dac8b2-2e15-4945-abef-ce98d248bb8e"
-)
+# Read in the raw voter statistics data. 
+readr::read_csv("inputs/data/raw_voter_statistics.csv")
 
-# write data
-write.csv (
-  x = raw_voting_locations,
-  file = "inputs/data/raw_voting_locations.csv"
-)
-
-#### Download data and write municipal election voter statistics data ####
-
-# download data
-raw_voter_statistics = download_data_from_opendatatoronto (
-  package_id = "7dc606ab-f042-4d90-99e2-9247cb5953d2",
-)
-
-# write data
-write.csv (
-  x = raw_voter_statistics,
-  file = "inputs/data/raw_voter_statistics.csv"
-)
+# Read in the raw voting locations data. 
+readr::read_csv("inputs/data/raw_voting_locations.csv")

### 2023-05-26 scripts/01-download_data.R
removed the source and need for the helper function

@@ -11,7 +11,6 @@ library(opendatatoronto)
 library(tidyverse)
 library(dplyr)
 library(readr)
-source("scripts/04-helper_functions.R")
 
 #### Download data and write municipal election polling location data ####
 

### 2023-05-26 scripts/04-helper_functions.R
Got my code to work in the "download" script rendering this script useless

@@ -1,18 +1,0 @@
-#### Preamble ####
-# Author: Inessa De Angelis
-# Date: 25 May 2023
-# Contact: inessa.deangelis@mail.utoronto.ca 
-# License: MIT
-# Pre-requisites: none
-
-#### Workspace setup ####
-
-library(opendatatoronto)
-
-## Download data from opendatatoronto
-download_data_from_opendatatoronto = function(package_id, resource_id) {
-  
-  raw_data = 
-    list_package_resources(package_id) |> # get resources for the package_id
-    get_resource() #download data
-}

### 2023-05-26 scripts/01-download_data.R
Minor edits to date and phrasing of section header

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Downloads and saves the data from OpenDataToronto portal
 # Author: Inessa De Angelis
-# Date: 25 May 2023
+# Date: 28 May 2023
 # Contact: inessa.deangelis@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: none
@@ -12,7 +12,7 @@ library(tidyverse)
 library(dplyr)
 library(readr)
 
-#### Download data and write municipal election polling location data ####
+#### Download data and write municipal election data ####
 
 # Read in the raw voter statistics data. 
 readr::read_csv("inputs/data/raw_voter_statistics.csv")

### 2023-05-26 scripts/00-simulate_data.R
Added "janitor" library to script. Amended a script, from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html to simulate fictional election turnouts by ward.

@@ -20,4 +20,8 @@ library(tidyverse)
 
 #### Simulate data ####
 
+simulated_data <-
+  tibble(
+  #use 1 through 25 to represent each ward
+  )
 

### 2023-05-28 scripts/01-download_data.R
Updated purpose and section headers

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from OpenDataToronto portal
+# Purpose: Downloads, saves, and reads data from the OpenDataToronto portal
 # Author: Inessa De Angelis
 # Date: 28 May 2023
 # Contact: inessa.deangelis@mail.utoronto.ca
@@ -13,6 +13,9 @@ library(dplyr)
 library(readr)
 
 #### Download data and write municipal election data ####
+
+
+#### Read municipal election data ####
 
 # Read in the raw voter statistics data. 
 readr::read_csv("inputs/data/raw_voter_statistics.csv")

### 2023-05-28 scripts/01-download_data.R
Using OpenDataToronto's package guide (https://cran.r-project.org/web/packages/opendatatoronto/opendatatoronto.pdf), I amended the code to directly download the voting locations data set from OpenDataToronto's website into this script.

@@ -13,7 +13,11 @@ library(dplyr)
 library(readr)
 
 #### Download data and write municipal election data ####
+# Code referenced from: https://cran.r-project.org/web/packages/opendatatoronto/opendatatoronto.pdf 
 
+list_package_resources("https://open.toronto.ca/dataset/elections-voting-locations/") %>%
+  head(1) %>%
+  get_resource()
 
 #### Read municipal election data ####
 

### 2023-05-28 scripts/01-download_data.R
Using the same code as the one I amended to download the voting locations data, I edited the code to download the voter statistics data, using the package resource id, instead of the URL.

@@ -12,12 +12,19 @@ library(tidyverse)
 library(dplyr)
 library(readr)
 
-#### Download data and write municipal election data ####
+#### Download voting locations data ####
 # Code referenced from: https://cran.r-project.org/web/packages/opendatatoronto/opendatatoronto.pdf 
 
 list_package_resources("https://open.toronto.ca/dataset/elections-voting-locations/") %>%
   head(1) %>%
   get_resource()
+
+#### Download voter statistics data ####
+# Code referenced from: https://cran.r-project.org/web/packages/opendatatoronto/opendatatoronto.pdf 
+
+list_package_resources("7dc606ab-f042-4d90-99e2-9247cb5953d2") %>%
+  head(1) %>%
+get_resource()
 
 #### Read municipal election data ####
 

### 2023-05-28 scripts/01-download_data.R
Minor edits to the formatting of "get_resource()" to see if it would properly download the data set

@@ -24,7 +24,7 @@ list_package_resources("https://open.toronto.ca/dataset/elections-voting-locatio
 
 list_package_resources("7dc606ab-f042-4d90-99e2-9247cb5953d2") %>%
   head(1) %>%
-get_resource()
+  get_resource()
 
 #### Read municipal election data ####
 

### 2023-05-28 scripts/01-download_data.R
Fixed issue with the voter statistics data not downloading directly from the portal into my script by specifying that we only want the 2022 data set. Code to specify this was referenced from: https://github.com/sharlagelfand/opendatatoronto/blob/main/vignettes/articles/multiple_resources_purrr.Rmd

@@ -20,9 +20,11 @@ list_package_resources("https://open.toronto.ca/dataset/elections-voting-locatio
   get_resource()
 
 #### Download voter statistics data ####
-# Code referenced from: https://cran.r-project.org/web/packages/opendatatoronto/opendatatoronto.pdf 
+# Code referenced from: https://cran.r-project.org/web/packages/opendatatoronto/opendatatoronto.pdf
+# & https://github.com/sharlagelfand/opendatatoronto/blob/main/vignettes/articles/multiple_resources_purrr.Rmd
 
 list_package_resources("7dc606ab-f042-4d90-99e2-9247cb5953d2") %>%
+  filter(grepl("2022", name)) %>%
   head(1) %>%
   get_resource()
 

### 2023-05-28 scripts/00-simulate_data.R
Line spacing.

@@ -112,3 +112,4 @@ data <- data.frame(
 # Barplot
 ggplot(data, aes(x=turnout, y=subdivisions)) + 
   geom_bar(stat = "identity") 
+

### 2023-05-28 scripts/00-simulate_data.R
Added header and began writing expectations for testing the simulated data

@@ -112,3 +112,8 @@ data <- data.frame(
 # Barplot
 ggplot(data, aes(x=turnout, y=subdivisions)) + 
   geom_bar(stat = "identity") 
+
+#### Test simulated data ####
+
+# Expectations #
+# There is no fewer than 1 ward and no more than 25 wards #

### 2023-05-28 scripts/02-data_cleaning.R
Updated script organization with re-naming of headers and sub-headers. Removed example code. Began writing code to select specific columns of interest from the voter statistics spreadsheet. Will return to this section of code to resolve errors.

@@ -10,39 +10,21 @@
 library(tidyverse)
 library(janitor)
 
-#### Clean data ####
+#### Read in raw data ####
+
 # Read in the raw voter statistics data. 
 readr::read_csv("inputs/data/raw_voter_statistics.csv")
 
 # Read in the raw voting locations data. 
 readr::read_csv("inputs/data/raw_voting_locations.csv")
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
+#### Data Cleaning ####
 
-#### Save data ####
-write_csv(cleaned_data, "outputs/data/analysis_data.csv")
+# select columns of interest #
+raw_voter_statistics |>
+  select(
+    ward,
+    sub,
+    total_eligible_electors,
+    number_voted,
+  )

### 2023-05-29 scripts/01-download_data.R
Updated all pipe code from %>% to |> to reflect newer code styling.

@@ -15,17 +15,17 @@ library(readr)
 #### Download voting locations data ####
 # Code referenced from: https://cran.r-project.org/web/packages/opendatatoronto/opendatatoronto.pdf 
 
-list_package_resources("https://open.toronto.ca/dataset/elections-voting-locations/") %>%
-  head(1) %>%
+list_package_resources("https://open.toronto.ca/dataset/elections-voting-locations/") |>
+  head(1) |>
   get_resource()
 
 #### Download voter statistics data ####
 # Code referenced from: https://cran.r-project.org/web/packages/opendatatoronto/opendatatoronto.pdf
 # & https://github.com/sharlagelfand/opendatatoronto/blob/main/vignettes/articles/multiple_resources_purrr.Rmd
 
-list_package_resources("7dc606ab-f042-4d90-99e2-9247cb5953d2") %>%
-  filter(grepl("2022", name)) %>%
-  head(1) %>%
+list_package_resources("7dc606ab-f042-4d90-99e2-9247cb5953d2") |>
+  filter(grepl("2022", name)) |>
+  head(1) |>
   get_resource()
 
 #### Read municipal election data ####

### 2023-05-29 scripts/02-data_cleaning.R
Added basic code, referring to https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html to clean the raw voter statistics data set, including adjusting column names.

@@ -18,10 +18,22 @@ readr::read_csv("inputs/data/raw_voter_statistics.csv")
 # Read in the raw voting locations data. 
 readr::read_csv("inputs/data/raw_voting_locations.csv")
 
-#### Data Cleaning ####
+#### Basic cleaning ####
+# based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
+raw_voter_statistics <-
+  read_csv(
+    file = "inputs/data/raw_voter_statistics.csv",
+    show_col_types = FALSE
+  )
+
+# Name organization#
+cleaned_voter_statistics <-
+  clean_names(raw_voter_statistics)
+
+head(cleaned_voter_statistics)
 
 # select columns of interest #
-raw_voter_statistics |>
+cleaned_voter_statistics <-
   select(
     ward,
     sub,

### 2023-05-29 scripts/02-data_cleaning.R
Added code, referencing: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html to rename the "%_voted" column to "percentrage_voted" because the percentage symbol at the start of the column name is causing R to say there is an unexpected input.

@@ -32,11 +32,22 @@ cleaned_voter_statistics <-
 
 head(cleaned_voter_statistics)
 
+# Rename % voted column #
+cleaned_voter_statistics <-
+  cleaned_voter_statistics |>
+  rename(
+    %_voted = percentage_voted,
+  )
+
+head(cleaned_voter_statistics)
+
 # select columns of interest #
 cleaned_voter_statistics <-
+  cleaned_voter_statistics |>
   select(
     ward,
     sub,
     total_eligible_electors,
     number_voted,
   )
+head(cleaned_voter_statistics)

### 2023-05-29 scripts/02-data_cleaning.R
Following the structure of the code I amended for the raw voter statistics data set, I added code to do basic cleaning for the raw voting locations data set, including simplifying column names and selecting specific columns of interest

@@ -18,7 +18,7 @@ readr::read_csv("inputs/data/raw_voter_statistics.csv")
 # Read in the raw voting locations data. 
 readr::read_csv("inputs/data/raw_voting_locations.csv")
 
-#### Basic cleaning ####
+#### Basic cleaning - voter statistics ####
 # based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
 raw_voter_statistics <-
   read_csv(
@@ -33,6 +33,7 @@ cleaned_voter_statistics <-
 head(cleaned_voter_statistics)
 
 # Rename % voted column #
+# based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
 cleaned_voter_statistics <-
   cleaned_voter_statistics |>
   rename(
@@ -51,3 +52,29 @@ cleaned_voter_statistics <-
     number_voted,
   )
 head(cleaned_voter_statistics)
+
+#### Basic cleaning - voting locations ####
+
+#### Basic cleaning ####
+# based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
+raw_voting_locations<-
+  read_csv(
+    file = "inputs/data/raw_voting_locations.csv",
+    show_col_types = FALSE
+  )
+
+# Name organization#
+cleaned_voting_locations <-
+  clean_names(raw_voting_locations)
+
+head(cleaned_voting_locations)
+
+# select columns of interest #
+cleaned_voting_locations <-
+  cleaned_voting_locations |>
+  select(
+    point_short_code,
+    point_long_code,
+    geometry
+  )
+head(cleaned_voting_locations)

### 2023-05-29 scripts/02-data_cleaning.R
Updated purpose to reflect the types of cleaning this script does for the raw data sets.

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Cleans the raw municipal election data by...
+# Purpose: Cleans the raw voter statistics and voting locations data by simplifying names and selecting the relevant columns from each data set
 # Author: Inessa De Angelis
 # Date: 25 May 2023
 # Contact: inessa.deangelis@mail.utoronto.ca 

### 2023-05-29 scripts/02-data_cleaning.R
Amended code from https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html to save the cleaned voting locations data set as a CSV file

@@ -26,7 +26,7 @@ raw_voter_statistics <-
     show_col_types = FALSE
   )
 
-# Name organization#
+# Name organization #
 cleaned_voter_statistics <-
   clean_names(raw_voter_statistics)
 
@@ -63,7 +63,7 @@ raw_voting_locations<-
     show_col_types = FALSE
   )
 
-# Name organization#
+# Name organization #
 cleaned_voting_locations <-
   clean_names(raw_voting_locations)
 

### 2023-05-29 scripts/00-simulate_data.R
Removed code to install tidyverse and janitor packages

@@ -11,6 +11,8 @@
 # Note: Only download the 2022 election data from the webpages
 
 #### Workspace setup ####
+# install.packages("tidyverse")
+# install.packages("janitor")
 library(tidyverse)
 library(janitor)
 
@@ -115,3 +117,7 @@ ggplot(data, aes(x=turnout, y=subdivisions)) +
 
 # Expectations #
 # There is no fewer than 1 ward and no more than 25 wards #
+simulate_data |>
+  runif(n = ward, min = 1, max = 25)
+  
+  

### 2023-05-29 scripts/02-data_cleaning.R
Amended code from https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html to save the cleaned voting locations data set as a CSV file

@@ -78,3 +78,10 @@ cleaned_voting_locations <-
     geometry
   )
 head(cleaned_voting_locations)
+
+# save cleaned voting locations data #
+# based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
+write_csv(
+  x = cleaned_voting_locations,
+  file = "cleaned_voting_locations.csv"
+)

### 2023-05-29 scripts/00-simulate_data.R
Removed code to install tidyverse and janitor packages

@@ -11,8 +11,6 @@
 # Note: Only download the 2022 election data from the webpages
 
 #### Workspace setup ####
-# install.packages("tidyverse")
-# install.packages("janitor")
 library(tidyverse)
 library(janitor)
 

### 2023-05-29 scripts/00-simulate_data.R
Updated purpose to better reflect what I am trying to simulate from the two data sets.

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Simulate the 2022 Toronto Municipal Election ward locations and results data sets 
+# Purpose: Simulate the 2022 Toronto Municipal Election voter statistics and voting locations data sets 
 # Author: Inessa De Angelis
 # Date: 28 May 2023 
 # Contact: inessa.deangelis@mail.utoronto.ca

### 2023-05-29 scripts/02-data_cleaning.R
Removed code to rename "% voted" column because previous code in the script to simplify all column names already changed the name to "percent_voted"

@@ -32,16 +32,6 @@ cleaned_voter_statistics <-
 
 head(cleaned_voter_statistics)
 
-# Rename % voted column #
-# based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
-cleaned_voter_statistics <-
-  cleaned_voter_statistics |>
-  rename(
-    %_voted = percentage_voted,
-  )
-
-head(cleaned_voter_statistics)
-
 # select columns of interest #
 cleaned_voter_statistics <-
   cleaned_voter_statistics |>
@@ -50,6 +40,7 @@ cleaned_voter_statistics <-
     sub,
     total_eligible_electors,
     number_voted,
+    percent_voted
   )
 head(cleaned_voter_statistics)
 

### 2023-05-29 scripts/02-data_cleaning.R
Using code I previously amended in this script, I updated the code to save the cleaned voter statistics data set, opting to save the cleaned data as a csv file.

@@ -44,6 +44,13 @@ cleaned_voter_statistics <-
   )
 head(cleaned_voter_statistics)
 
+# save cleaned voter statistics data #
+# based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
+write_csv(
+  x = cleaned_voter_statistics,
+  file = "cleaned_voter_statistics.csv"
+)
+
 #### Basic cleaning - voting locations ####
 
 #### Basic cleaning ####

### 2023-05-30 scripts/01-download_data.R
Amended the script to save the downloaded raw voting locations and voter statistics data

@@ -19,6 +19,13 @@ list_package_resources("https://open.toronto.ca/dataset/elections-voting-locatio
   head(1) |>
   get_resource()
 
+# save raw voting locations data #
+# based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
+write_csv(
+  x = raw_voting_locations,
+  file = "raw_voting_locations.csv"
+)
+
 #### Download voter statistics data ####
 # Code referenced from: https://cran.r-project.org/web/packages/opendatatoronto/opendatatoronto.pdf
 # & https://github.com/sharlagelfand/opendatatoronto/blob/main/vignettes/articles/multiple_resources_purrr.Rmd
@@ -27,6 +34,13 @@ list_package_resources("7dc606ab-f042-4d90-99e2-9247cb5953d2") |>
   filter(grepl("2022", name)) |>
   head(1) |>
   get_resource()
+
+# save raw voter statistics data #
+# based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
+write_csv(
+  x = raw_voter_statistics,
+  file = "raw_voter_statistics.csv"
+)
 
 #### Read municipal election data ####
 

### 2023-05-30 scripts/00-simulate_data.R
Amended code from: https://github.com/christina-wei/INF3104-1-Covid-Clinics/blob/main/scripts/00-simulation.R to test if there are no more than 25 wards

@@ -111,11 +111,14 @@ data <- data.frame(
 ggplot(data, aes(x=turnout, y=subdivisions)) + 
   geom_bar(stat = "identity") 
 
-#### Test simulated data ####
+#### Data Validation ####
 
-# Expectations #
-# There is no fewer than 1 ward and no more than 25 wards #
+# There are no more than 25 wards #
+# Referenced code from: https://github.com/christina-wei/INF3104-1-Covid-Clinics/blob/main/scripts/00-simulation.R
 simulate_data |>
-  runif(n = ward, min = 1, max = 25)
+  group_by(ward) |>
+  count() |>
+  filter(n > ward) |>
+  sum() == 0
   
   

### 2023-05-30 scripts/00-simulate_data.R
Amended code previously used in this script to check simulated voter turnout by ward

@@ -113,12 +113,19 @@ ggplot(data, aes(x=turnout, y=subdivisions)) +
 
 #### Data Validation ####
 
-# There are no more than 25 wards #
+# Check that there are no more than 25 wards #
 # Referenced code from: https://github.com/christina-wei/INF3104-1-Covid-Clinics/blob/main/scripts/00-simulation.R
 simulate_data |>
   group_by(ward) |>
   count() |>
   filter(n > ward) |>
   sum() == 0
-  
+
+# Check that voter turnout per ward is between 19% and 44% #
+simulate_data |>
+  group_by(turnout) |>
+  count() |>
+  filter(n > turnout)
+sum() == 0
+
   

### 2023-05-30 scripts/00-simulate_data.R
Edited script to check that the simulated voter turnout is between 19% and 44%.
Also added script, referencing: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html to check that there are no fewer than 1 ward and no more than 25.

@@ -113,6 +113,11 @@ ggplot(data, aes(x=turnout, y=subdivisions)) +
 
 #### Data Validation ####
 
+# Check that there are between 1 and 25 wards #
+# Referenced code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
+simulate_data$ward |> min() == 1
+simulate_data$ward |> max() == 25
+
 # Check that there are no more than 25 wards #
 # Referenced code from: https://github.com/christina-wei/INF3104-1-Covid-Clinics/blob/main/scripts/00-simulation.R
 simulate_data |>
@@ -122,10 +127,5 @@ simulate_data |>
   sum() == 0
 
 # Check that voter turnout per ward is between 19% and 44% #
-simulate_data |>
-  group_by(turnout) |>
-  count() |>
-  filter(n > turnout)
-sum() == 0
-
-  
+simulate_data$turnout |> min() == 19
+simulate_data$turnout |> max() == 44

### 2023-05-30 scripts/00-simulate_data.R
Edited code previously used in this script to check that the number of simulated subdivisions per ward is between 50 and 75.

@@ -129,3 +129,10 @@ simulate_data |>
 # Check that voter turnout per ward is between 19% and 44% #
 simulate_data$turnout |> min() == 19
 simulate_data$turnout |> max() == 44
+
+# Check that the number of subdivisions per ward is between 50 and 75 #
+simulate_data |>
+  group_by(number_subdivision) |>
+  count() |>
+  filter(n > number_subdivision) |>
+  sum() == 0

### 2023-05-30 outputs/paper/paper.qmd
Set up work space and loaded tidyverse package. Began writing code to read in the cleaned voter statistics and voting locations data sets.
Updating file names for the code to read in the cleaned data sets.
Adjusting the script to read in the two cleaned data sets
Removed existing code to read in cleaned data because it had errors
Added "here" package to help read in the cleaned data. Wrote code to read in the cleaned voter statistics and voting locations data sets.

@@
 ```{r}
+#| message: false
+#| echo: false
 
+#### Workspace set-up ####
+
+## Load packages
+library(tidyverse)
+library(knitr)
+library(here)
+
+#read in cleaned data #
+cleaned_voter_statistics <- read.csv(here::here("inputs/data/cleaned_voter_statistics.csv"))
+show_col_types = FALSE
+
+cleaned_voting_locations <- read.csv(here::here("inputs/data/cleaned_voting_locations.csv"))
+show_col_types = FALSE
 ```
+
+## Introduction

### 2023-05-30 scripts/03-test_data.R
Deleting separate testing script as it is not required for this paper

@@ -1,14 +1,0 @@
-#### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Inessa De Angelis
-# Date: 22 May 2023
-# Contact: inessa.deangelis@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: none
-
-
-#### Workspace setup ####
-library(tidyverse)
-# [...UPDATE THIS...]
-
-#### Test data ####

### 2023-05-30 outputs/paper/paper.qmd
Added introduction paragaphs and began formatting citations

@@
 author: "Inessa De Angelis"
-thanks: "Code and data from this analysis are available at: https://github.com/InessaDeAngelis/Toronto_Elections."
+thanks: "Code and data from this analysis are available at: https://github.com/InessaDeAngelis/Toronto_Elections"
 date: "today"
@@
 ## Introduction
+As mandated by the Government of Ontario, residents of the City of Toronto went to the polls on October 24, 2022 to elect a mayor, councillors, and school board trustees.
+Then-sitting Mayor, John Tory sought re-election, along with a number of incumbent city councillors.
+Coming out of the height of the Covid-19 pandemic, the 2022 election largely upheld the status quo and did not feature ambitious policy platforms.
+Voter turnout across the city was 29.7% - the lowest in the city’s history since amalgamation in 1997, with turnout ranging on a ward-by-ward basis from 22.1% to 38.5% (Warren, 2022; Marshall, 2023).
+
+Individual turnout by ward largely depended on local candidates and the accessibility of subdivisions to cast a ballot.
+The City of Toronto @ElectionDictionary defines a ward as “a geographical area represented by a member of Council.” Following Premier Doug Ford’s decision in the summer of 2018 to cut down the number of wards in the city to match federal and provincial riding boundaries, there are currently 25 wards in the City of Toronto (Lucas & McGregor, 2021).
+Subdivisions are defined by the City of Toronto [@VoterStatistics] as “....
+geographical area[s within a ward] designated by the City Clerk.” Previous studies concentrating on large cities in the United States, including Atlanta revealed that having polling locations in close proximity to a voter’s home bolsters turnout and even minor changes in placement of a polling location can have significant impact on a voter’s decision to cast a ballot (Haspel & Knotts, 2005).
+There has been little research into the impact of polling location placement in relation to election turnout, especially in Canadian cities like Toronto and this paper will contribute to the investigation of this phenomenon.
+
+To examine the placement of subdivisions on voter turnout, this paper is organized into the following sections: Data, Results, Discussion, and Conclusion.
+In the Data section, I discuss the nature of the spreadsheets obtained through the City of Toronto’s OpenDataToronto Library [@rOpenDataToronto] and the steps I took to clean and analyze the data.
+The Results section highlights trends found during the analysis process, while the Discussion section further evaluates the trends and presents insight.
+Lastly, the Conclusion section summarizes the main findings from this paper.

### 2023-05-30 outputs/paper/paper.qmd
Added data section and elections voting locations data sub-section and began formatting citations for both sections

@@
 Lastly, the Conclusion section summarizes the main findings from this paper.
+
+## Data
+The data utilized throughout this paper was obtained through the City of Toronto’s OpenDataToronto Library [@rOpenDataToronto].
+The two different data sets used are: “Elections Voting Locations” (Data, 2022) and “Elections - Voter Statistics” (Data, 2023).
+Data was collected and analyzed using the statistical programming software R (R Core Team, 2022), with additional support from from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], and `janitor` [@rJanitor].
+A further discussion of the data collection, cleaning, and analysis process can be found later on in this paper.
+
+### Elections Voting Locations
+This dataset, published by the City Clerk's Office [@VotingLocations] highlights all polling locations and subdivisions, on a ward-by-ward basis from the 2022 municipal election.
+This data was last refreshed on November 2, 2022 and captured for this paper on May 21, 2023.
+The polling locations account for Advance Vote, election day, and Mail In Voting.
+One subdivision in a ward is equivalent to one polling location, based on geography.
+In each ward, subdivision 97 accounts for Mail In Voting, while subdivisions 98 and 99 are designated Advance Vote locations.
+Voters can request a Mail In ballot, go to one of the two Advance Vote locations during the designated days, or vote at their assigned subdivision on election day (for example: ward 11, subdivision 11).

### 2023-05-30 outputs/paper/paper.qmd
Amended code to provide a sample of cleaned data from the voter statistics data set. Also added a paragraph outlining the voter statistics data set.

@@
 Coming out of the height of the Covid-19 pandemic, the 2022 election largely upheld the status quo and did not feature ambitious policy platforms.
-Voter turnout across the city was 29.7% - the lowest in the city’s history since amalgamation in 1997, with turnout ranging on a ward-by-ward basis from 22.1% to 38.5% (Warren, 2022; Marshall, 2023).
+Voter turnout across the city was 29.7% - the lowest in the city’s history since amalgamation in 1997, with turnout ranging on a ward-by-ward basis from 22.1% to 38.5% [@Turnout22; @DismalTurnout].
 
 Individual turnout by ward largely depended on local candidates and the accessibility of subdivisions to cast a ballot.
-The City of Toronto @ElectionDictionary defines a ward as “a geographical area represented by a member of Council.” Following Premier Doug Ford’s decision in the summer of 2018 to cut down the number of wards in the city to match federal and provincial riding boundaries, there are currently 25 wards in the City of Toronto (Lucas & McGregor, 2021).
+The City of Toronto @ElectionDictionary defines a ward as “a geographical area represented by a member of Council.” Following Premier Doug Ford’s decision in the summer of 2018 to cut down the number of wards in the city to match federal and provincial riding boundaries, there are currently 25 wards in the City of Toronto [@BigCityElxns].
 Subdivisions are defined by the City of Toronto [@VoterStatistics] as “....
-geographical area[s within a ward] designated by the City Clerk.” Previous studies concentrating on large cities in the United States, including Atlanta revealed that having polling locations in close proximity to a voter’s home bolsters turnout and even minor changes in placement of a polling location can have significant impact on a voter’s decision to cast a ballot (Haspel & Knotts, 2005).
+geographical area[s within a ward] designated by the City Clerk.” Previous studies concentrating on large cities in the United States, including Atlanta revealed that having polling locations in close proximity to a voter’s home bolsters turnout and even minor changes in placement of a polling location can have significant impact on a voter’s decision to cast a ballot [@Locationx3].
 There has been little research into the impact of polling location placement in relation to election turnout, especially in Canadian cities like Toronto and this paper will contribute to the investigation of this phenomenon.
@@
 The data utilized throughout this paper was obtained through the City of Toronto’s OpenDataToronto Library [@rOpenDataToronto].
-The two different data sets used are: “Elections Voting Locations” (Data, 2022) and “Elections - Voter Statistics” (Data, 2023).
+The two different data sets used are: “Elections Voting Locations” [@VotingLocations] and “Elections - Voter Statistics” [@VoterStatistics].
 Data was collected and analyzed using the statistical programming software R (R Core Team, 2022), with additional support from from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], and `janitor` [@rJanitor].
@@
 
-### Elections Voting Locations
+### Election Voting Locations
 This dataset, published by the City Clerk's Office [@VotingLocations] highlights all polling locations and subdivisions, on a ward-by-ward basis from the 2022 municipal election.
@@
 Voters can request a Mail In ballot, go to one of the two Advance Vote locations during the designated days, or vote at their assigned subdivision on election day (for example: ward 11, subdivision 11).
+
+### Election Voter Statistics
+This dataset, published by the City Clerk's Office [@VoterStatistics] comprises the number of eligible voters and corrections/additions to the Voter’s List for the 2022 municipal election.
+The dataset is broken down by ward and voting subdivision, including voting for school board trustees.
+This data set was last refreshed on February 7, 2023 and captured for this paper on May 21, 2023.
+This data set outlines the total number of eligible electors in Toronto, by ward, subdivision, and the entire city.
+
+Upon analysis, it was discovered that this data set included columns and data beyond the scope of this paper.
+I conducted basic cleaning of the data set to simplify column names and eliminate additional information (see tbl-cleaned_voter_statistics).
+
+```{r}
+#| message: false
+#| echo: false
+#| label: tbl-cleaned_voter_statistics
+#| tbl-cap: Sample of Cleaned Voter Statistics Data
+
+# code referenced from: https://github.com/christina-wei/INF3104-1-Covid-Clinics/blob/main/outputs/paper/covid_clinics.qmd
+head (cleaned_voter_statistics, 5) |>
+  kable(
+    col.names = c("ward", "sub", "total_eligible_electors",	"number_voted",	"percent_voted"),
+  booktabs = TRUE
+  )
+
+```

### 2023-05-30 outputs/paper/paper.qmd
Made minor edits to the title and formatting of Table 1.

@@
 ---
-title: "Do More Polling Locations in Wards Increase Turnout? An Analysis of the 2022 Toronto Municipal Election"
+title: "Do More Subdivisions in Wards Increase Turnout? An Analysis of the 2022 Toronto Municipal Election"
 author: "Inessa De Angelis"
@@
 
-To examine the placement of subdivisions on voter turnout, this paper is organized into the following sections: Data, Results, Discussion, and Conclusion.
+To examine the impact of subdivision placement on voter turnout, this paper is organized into the following sections: Data, Results, Discussion, and Conclusion.
 In the Data section, I discuss the nature of the spreadsheets obtained through the City of Toronto’s OpenDataToronto Library [@rOpenDataToronto] and the steps I took to clean and analyze the data.
@@
 Upon analysis, it was discovered that this data set included columns and data beyond the scope of this paper.
-I conducted basic cleaning of the data set to simplify column names and eliminate additional information (see tbl-cleaned_voter_statistics).
+I conducted basic cleaning of the data set to simplify column names and eliminate additional information (see @tbl-cleaned_voter_statistics).
 
@@
   kable(
-    col.names = c("ward", "sub", "total_eligible_electors",	"number_voted",	"percent_voted"),
+    col.names = c("Ward", "Sub", "Total Eligible Electors",	"Number Voted",	"Percent Voted"),
   booktabs = TRUE

### 2023-05-30 scripts/02-data_cleaning.R
Amended the script to include the cleaned voting locations file which includes full address information for each subdivision

@@ -73,6 +73,7 @@ cleaned_voting_locations <-
   select(
     point_short_code,
     point_long_code,
+    address_full,
     geometry
   )
 head(cleaned_voting_locations)

### 2023-05-30 outputs/paper/paper.qmd
Installed sf package for upcoming mapping of subdivision data.
Also, using example code referenced from 2 places, I put together a table using cleaned voting locations data, which shows subdivision, ward/subdivision numbers combined, and subdivision addresses.

@@
 library(here)
+library(sf)
 
@@
 In each ward, subdivision 97 accounts for Mail In Voting, while subdivisions 98 and 99 are designated Advance Vote locations.
-Voters can request a Mail In ballot, go to one of the two Advance Vote locations during the designated days, or vote at their assigned subdivision on election day (for example: ward 11, subdivision 11).
+Voters can request a Mail In ballot, go to one of the two Advance Vote locations during the designated days, or vote at their assigned subdivision on election day (for example: Ward 11, Subdivision 11).
+
+```{r}
+#| message: false
+#| echo: false
+#| label: tbl-cleaned_voting_location
+#| tbl-cap: Sample of Cleaned Voting Locations Data
+
+# code referenced from: https://github.com/christina-wei/INF3104-1-Covid-Clinics/blob/main/outputs/paper/covid_clinics.qmd
+# & https://tellingstorieswithdata.com/20-r_essentials.html 
+head (cleaned_voting_locations, 5) |>
+  select(point_short_code, point_long_code, address_full) |>
+  kable(
+    col.names = c("Subdivision", "Ward & Subdivision", "Address"),
+  booktabs = TRUE
+  )
+
+```
 
@@
 
-# code referenced from: https://github.com/christina-wei/INF3104-1-Covid-Clinics/blob/main/outputs/paper/covid_clinics.qmd
 head (cleaned_voter_statistics, 5) |>

### 2023-05-30 outputs/paper/paper.qmd
Added ggplot2 and ggspatial packages and began writing code to map the locations of subdivisions

@@
 library(sf)
+library(ggplot2)
+library(ggspatial)
 
@@
 ```
+
+### Results
+
+```{r}
+#| message: false
+#| echo: false
+
+# code referenced from: https://sharlagelfand.github.io/opendatatoronto/articles/articles/spatial_data.html
+cleaned_voting_locations |>
+  c("address_full", "geometry")
+ggplot() +
+  geom_sf(data = address) +
+  geom_sf(data = geometry) +
+  theme_void()
+
+```

### 2023-05-31 scripts/02-data_cleaning.R
Adding code to summarize the number of subdivisions per ward

@@ -84,3 +84,10 @@ write_csv(
   x = cleaned_voting_locations,
   file = "cleaned_voting_locations.csv"
 )
+
+# count number of subdivisions per ward #
+# based on code from: https://tellingstorieswithdata.com/20-r_essentials.html 
+cleaned_voter_statistics |>
+  filter("ward" == 1) |>
+summary("sub")
+  head(cleaned_voter_statistics)

### 2023-05-31 scripts/02-data_cleaning.R
continued trying code to summarize number of subdivisions per ward

@@ -9,6 +9,7 @@
 #### Workspace setup ####
 library(tidyverse)
 library(janitor)
+library(dplyr)
 
 #### Read in raw data ####
 
@@ -88,6 +89,8 @@ write_csv(
 # count number of subdivisions per ward #
 # based on code from: https://tellingstorieswithdata.com/20-r_essentials.html 
 cleaned_voter_statistics |>
-  filter("ward" == 1) |>
-summary("sub")
-  head(cleaned_voter_statistics)
+  filter(ward == 1) |>
+  summary("sub")
+head(cleaned_voter_statistics)
+
+  

### 2023-05-31 scripts/02-data_cleaning.R
amended script to summarise the number of subdivisions by ward

@@ -88,9 +88,8 @@ write_csv(
 
 # count number of subdivisions per ward #
 # based on code from: https://tellingstorieswithdata.com/20-r_essentials.html 
-cleaned_voter_statistics |>
-  filter(ward == 1) |>
-  summary("sub")
-head(cleaned_voter_statistics)
 
+cleaned_voter_statistics |>
+summarise(sub = n(),
+          .by = ward)
   

### 2023-05-31 scripts/02-data_cleaning.R
amending script to summarize number of subdivisions by ward

@@ -89,7 +89,13 @@ write_csv(
 # count number of subdivisions per ward #
 # based on code from: https://tellingstorieswithdata.com/20-r_essentials.html 
 
+summarized_voter_statistics = 
+  cleaned_voter_statistics |>
+  group_by(ward) |>
+  count(sub)
+head(summarized_voter_statistics)
+  
 cleaned_voter_statistics |>
 summarise(sub = n(),
           .by = ward)
-  
+

### 2023-05-31 outputs/paper/paper.qmd
trying to visualize number of subdivisions by ward

@@
 library(knitr)
+library(kableExtra)
 library(here)
@@
 library(ggspatial)
+library(lubridate)
 
@@
 ```
+
+```{r}
+#| message: false
+#| echo: false
+
+# summarize ward data by number of subdivisions
+
+head (cleaned_voter_statistics, 5) |>
+ select(ward, sub, percent_voted) |>
+  arrange (ward) |>
+   kable(
+    col.names = c("Ward", "Sub", "Percent Voted"),
+  booktabs = TRUE
+  )
+
+cleaned_voter_statistics |>
+  slice(2:53)
+
+  filter("ward" == 1) |>
+  select(ward, sub)
+
+```

### 2023-05-31 scripts/02-data_cleaning.R
Added code to summarize the number of subdivisions on a ward-by-ward basis.

@@ -90,12 +90,21 @@ write_csv(
 # based on code from: https://tellingstorieswithdata.com/20-r_essentials.html 
 
 summarized_voter_statistics = 
-  cleaned_voter_statistics |>
-  group_by(ward) |>
-  count(sub)
-head(summarized_voter_statistics)
-  
 cleaned_voter_statistics |>
-summarise(sub = n(),
-          .by = ward)
+  slice(2:52) |>
+ reframe(sub) |>
+  summarise(
+    total = max(sub, na.rm = TRUE)
+  )
+summarized_voter_statistics
+
+
+summarized_voter_statistics = 
+  cleaned_voter_statistics |>
+  slice(58:125) |>
+  reframe(sub) |>
+  summarise(
+    total = max(sub, na.rm = TRUE)
+  )
+summarized_voter_statistics
 

### 2023-05-31 scripts/02-data_cleaning.R
Updated code I wrote to summarize the number of subdivisions in ward 1.

@@ -89,22 +89,11 @@ write_csv(
 # count number of subdivisions per ward #
 # based on code from: https://tellingstorieswithdata.com/20-r_essentials.html 
 
-summarized_voter_statistics = 
-cleaned_voter_statistics |>
-  slice(2:52) |>
- reframe(sub) |>
-  summarise(
-    total = max(sub, na.rm = TRUE)
-  )
-summarized_voter_statistics
-
-
+#Ward 1
 summarized_voter_statistics = 
   cleaned_voter_statistics |>
-  slice(58:125) |>
-  reframe(sub) |>
-  summarise(
-    total = max(sub, na.rm = TRUE)
-  )
+  slice(2:52) |>
+  count(
+    total = max(sub, na.rm = TRUE),
 summarized_voter_statistics
-
+)

### 2023-05-31 outputs/paper/paper.qmd
Updated code to count subdivisions by ward

@@
 head (cleaned_voter_statistics, 5) |>
- select(ward, sub, percent_voted) |>
+ select(ward, sub) |>
   arrange (ward) |>
    kable(
-    col.names = c("Ward", "Sub", "Percent Voted"),
+    col.names = c("Ward", "Sub"),
   booktabs = TRUE
@@
 cleaned_voter_statistics |>
-  slice(2:53)
-
-  filter("ward" == 1) |>
-  select(ward, sub)
+  slice(2:52) |>
+  count(
+    total = max(sub, na.rm = TRUE)
+  )
 

### 2023-06-01 outputs/paper/paper.qmd
Edited typo

@@
 The two different data sets used are: “Elections Voting Locations” [@VotingLocations] and “Elections - Voter Statistics” [@VoterStatistics].
-Data was collected and analyzed using the statistical programming software R (R Core Team, 2022), with additional support from from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], and `janitor` [@rJanitor].
+Data was collected and analyzed using the statistical programming software R (R Core Team, 2022), with additional support from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], and `janitor` [@rJanitor].
 A further discussion of the data collection, cleaning, and analysis process can be found later on in this paper.

### 2023-06-01 outputs/paper/paper.qmd
Added additional section headers and wrote a further explanation about the voter statistics data set

@@
 
+For the purpose of this paper, I did not include the Advance Vote and Mail In Voting subdivisions when counting the total number of subdivisions per ward.
+Firstly, there are only two Advance Vote locations distributed across every ward, meaning residents have to be motivated to travel farther away from their home to cast their ballot.
+Secondly, Mail In Voting requires applying online through Toronto Elections and then returning the completed Mail In Voting package by dropping it in a Canada Post mailbox or hand delivering it to one of the Toronto Elections drop boxes at select locations.
+Voters need to be more organized and motivated to vote in advance of election day.
+
 ```{r}
@@
 ```
+### Discussion
+
+### Conclusion
+
+### References

### 2023-06-01 outputs/paper/paper.qmd
Added code to generate table of contents
Removed packages and code to plot spatial data

@@
 format: pdf
+toc: true
 number-sections: true
@@
 library(here)
-library(sf)
 library(ggplot2)
-library(ggspatial)
 library(lubridate)
@@
 
-### Results
+## Results
 
@@
 
-# code referenced from: https://sharlagelfand.github.io/opendatatoronto/articles/articles/spatial_data.html
-cleaned_voting_locations |>
-  c("address_full", "geometry")
-ggplot() +
-  geom_sf(data = address) +
-  geom_sf(data = geometry) +
-  theme_void()
+#code referenced from: https://github.com/christina-wei/INF3104-1-Covid-Clinics/blob/main/outputs/paper/covid_clinics.qmd
+cleaned_voter_statistics |>
+  ggplot(aes(x = ward), "-", sub, fill = location_type) + 
+  geom_bar() + #bar graph
+  theme_minimal() + #neater theme
+  theme(axis.text.x = element_text(angle=90, size = 6)) + #rotate x-axis & smaller font
+  labs(
+    title = "Number of Subdivisions Per Ward in Toronto",
+    x = "Ward",
+    y = "Number of Subdivisions",
+    fill = "Type"
+  )  #a
 
@@
 
-head (cleaned_voter_statistics, 5) |>
+head (cleaned_voter_statistics, 60) |>
  select(ward, sub) |>
@@
 ```
-### Discussion
+## Discussion
 
-### Conclusion
+## Conclusion
 
-### References
+## References

### 2023-06-01 scripts/02-data_cleaning.R
added additional script which could also count the number of subdivisions per ward, using the filter() and count() functions

@@ -97,3 +97,19 @@ summarized_voter_statistics =
     total = max(sub, na.rm = TRUE),
 summarized_voter_statistics
 )
+
+# Filter Ward 1 only
+summarized_voter_statistics = 
+  cleaned_voter_statistics |>
+  filter(ward == 1) 
+  summarized_voter_statistics
+  
+
+#Filter Ward 1 only and count number of subdivisions 
+summarized_voter_statistics = 
+  cleaned_voter_statistics |>
+  filter(ward == 24) |>
+  count(
+    total = max(sub, na.rm = TRUE)
+  )
+summarized_voter_statistics

### 2023-06-01 outputs/paper/paper.qmd
Updated citations for packages used

@@
 The two different data sets used are: “Elections Voting Locations” [@VotingLocations] and “Elections - Voter Statistics” [@VoterStatistics].
-Data was collected and analyzed using the statistical programming software R (R Core Team, 2022), with additional support from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], and `janitor` [@rJanitor].
+Data was collected and analyzed using the statistical programming software R (@r), with additional support from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], and `here` [@rHere].
 A further discussion of the data collection, cleaning, and analysis process can be found later on in this paper.

### 2023-06-02 scripts/02-data_cleaning.R
Updating script to filter by ward and count number of subdivisions per ward

@@ -89,27 +89,21 @@ write_csv(
 # count number of subdivisions per ward #
 # based on code from: https://tellingstorieswithdata.com/20-r_essentials.html 
 
-#Ward 1
+#Filter by ward and count number of subdivisions 
 summarized_voter_statistics = 
   cleaned_voter_statistics |>
-  slice(2:52) |>
+  group_by(ward) |>
   count(
-    total = max(sub, na.rm = TRUE),
+    total = max(sub, na.rm = TRUE) 
+  ) |>
+  arrange(desc(ward)) |>
+  arrange(desc(n))
 summarized_voter_statistics
+
+
+# save summarized voter statistics data #
+# based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
+write_csv(
+  x = summarized_voter_statistics,
+  file = "summarized_voter_statistics.csv"
 )
-
-# Filter Ward 1 only
-summarized_voter_statistics = 
-  cleaned_voter_statistics |>
-  filter(ward == 1) 
-  summarized_voter_statistics
-  
-
-#Filter Ward 1 only and count number of subdivisions 
-summarized_voter_statistics = 
-  cleaned_voter_statistics |>
-  filter(ward == 24) |>
-  count(
-    total = max(sub, na.rm = TRUE)
-  )
-summarized_voter_statistics

### 2023-06-02 scripts/02-data_cleaning.R
Added code to arrange by number and select only the ward and total number of subdivisions

@@ -97,7 +97,8 @@ summarized_voter_statistics =
     total = max(sub, na.rm = TRUE) 
   ) |>
   arrange(desc(ward)) |>
-  arrange(desc(n))
+  arrange(desc(n)) |>
+  select(ward, n)
 summarized_voter_statistics
 
 

### 2023-06-02 scripts/00-simulate_data.R
Removed reference to the voting locations data set

@@ -1,14 +1,13 @@
 #### Preamble ####
-# Purpose: Simulate the 2022 Toronto Municipal Election voter statistics and voting locations data sets 
+# Purpose: Simulate the 2022 Toronto Municipal Election voter statistics data set
 # Author: Inessa De Angelis
 # Date: 28 May 2023 
 # Contact: inessa.deangelis@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: none
 # Data sets:
-## - https://open.toronto.ca/dataset/elections-voting-locations/ 
 ## - https://open.toronto.ca/dataset/elections-voter-statistics/ 
-# Note: Only download the 2022 election data from the webpages
+# Note: Only download the 2022 election data from the webpage
 
 #### Workspace setup ####
 library(tidyverse)

### 2023-06-02 scripts/02-data_cleaning.R
Removed script to clean voting locations data set as it will not be used in this paper

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Cleans the raw voter statistics and voting locations data by simplifying names and selecting the relevant columns from each data set
+# Purpose: Cleans the raw voter statistics by simplifying names and selecting the relevant columns from the data set
 # Author: Inessa De Angelis
 # Date: 25 May 2023
 # Contact: inessa.deangelis@mail.utoronto.ca 
@@ -15,9 +15,6 @@ library(dplyr)
 
 # Read in the raw voter statistics data. 
 readr::read_csv("inputs/data/raw_voter_statistics.csv")
-
-# Read in the raw voting locations data. 
-readr::read_csv("inputs/data/raw_voting_locations.csv")
 
 #### Basic cleaning - voter statistics ####
 # based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
@@ -52,40 +49,6 @@ write_csv(
   file = "cleaned_voter_statistics.csv"
 )
 
-#### Basic cleaning - voting locations ####
-
-#### Basic cleaning ####
-# based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
-raw_voting_locations<-
-  read_csv(
-    file = "inputs/data/raw_voting_locations.csv",
-    show_col_types = FALSE
-  )
-
-# Name organization #
-cleaned_voting_locations <-
-  clean_names(raw_voting_locations)
-
-head(cleaned_voting_locations)
-
-# select columns of interest #
-cleaned_voting_locations <-
-  cleaned_voting_locations |>
-  select(
-    point_short_code,
-    point_long_code,
-    address_full,
-    geometry
-  )
-head(cleaned_voting_locations)
-
-# save cleaned voting locations data #
-# based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
-write_csv(
-  x = cleaned_voting_locations,
-  file = "cleaned_voting_locations.csv"
-)
-
 # count number of subdivisions per ward #
 # based on code from: https://tellingstorieswithdata.com/20-r_essentials.html 
 
@@ -98,9 +61,14 @@ summarized_voter_statistics =
   ) |>
   arrange(desc(ward)) |>
   arrange(desc(n)) |>
-  select(ward, n)
+  select(ward, n) 
 summarized_voter_statistics
 
+#add percent voted to summarized voter statistics
+summarized_voter_statistics <-
[…]

### 2023-06-02 scripts/01-download_data.R
Removed script to download, save, and read voting locations data

@@ -12,20 +12,6 @@ library(tidyverse)
 library(dplyr)
 library(readr)
 
-#### Download voting locations data ####
-# Code referenced from: https://cran.r-project.org/web/packages/opendatatoronto/opendatatoronto.pdf 
-
-list_package_resources("https://open.toronto.ca/dataset/elections-voting-locations/") |>
-  head(1) |>
-  get_resource()
-
-# save raw voting locations data #
-# based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
-write_csv(
-  x = raw_voting_locations,
-  file = "raw_voting_locations.csv"
-)
-
 #### Download voter statistics data ####
 # Code referenced from: https://cran.r-project.org/web/packages/opendatatoronto/opendatatoronto.pdf
 # & https://github.com/sharlagelfand/opendatatoronto/blob/main/vignettes/articles/multiple_resources_purrr.Rmd
@@ -46,6 +32,3 @@ write_csv(
 
 # Read in the raw voter statistics data. 
 readr::read_csv("inputs/data/raw_voter_statistics.csv")
-
-# Read in the raw voting locations data. 
-readr::read_csv("inputs/data/raw_voting_locations.csv")

### 2023-06-02 scripts/02-data_cleaning.R
Added data validation based on script I wrote in the simulation section

@@ -55,19 +55,13 @@ write_csv(
 #Filter by ward and count number of subdivisions 
 summarized_voter_statistics = 
   cleaned_voter_statistics |>
-  group_by(ward) |>
+ group_by(ward) |>
   count(
     total = max(sub, na.rm = TRUE) 
   ) |>
   arrange(desc(ward)) |>
   arrange(desc(n)) |>
-  select(ward, n) 
-summarized_voter_statistics
-
-#add percent voted to summarized voter statistics
-summarized_voter_statistics <-
- summarized_voter_statistics |>
-  left_join(cleaned_voter_statistics, by = "percent_voted")
+  select(ward, n)
 summarized_voter_statistics
 
 # save summarized voter statistics data #
@@ -76,3 +70,14 @@ write_csv(
   x = summarized_voter_statistics,
   file = "summarized_voter_statistics.csv"
 )
+
+#### Data validation ####
+
+# Check that there are between 1 and 25 wards #
+# Referenced code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
+summarized_voter_statistics$ward |> min() == 1
+summarized_voter_statistics$ward |> max() == 25
+
+# Check that number of subdivisions per ward is between 1 and 94
+summarized_voter_statistics$n |> min() == 1
+summarized_voter_statistics$n |> max() == 94

### 2023-06-02 outputs/paper/paper.qmd
Removed voting locations portion

@@
 show_col_types = FALSE
-
-cleaned_voting_locations <- read.csv(here::here("inputs/data/cleaned_voting_locations.csv"))
-show_col_types = FALSE
 ```
@@
 
-### Election Voting Locations
-This dataset, published by the City Clerk's Office [@VotingLocations] highlights all polling locations and subdivisions, on a ward-by-ward basis from the 2022 municipal election.
-This data was last refreshed on November 2, 2022 and captured for this paper on May 21, 2023.
-The polling locations account for Advance Vote, election day, and Mail In Voting.
-One subdivision in a ward is equivalent to one polling location, based on geography.
-In each ward, subdivision 97 accounts for Mail In Voting, while subdivisions 98 and 99 are designated Advance Vote locations.
-Voters can request a Mail In ballot, go to one of the two Advance Vote locations during the designated days, or vote at their assigned subdivision on election day (for example: Ward 11, Subdivision 11).
-
-```{r}
-#| message: false
-#| echo: false
-#| label: tbl-cleaned_voting_location
-#| tbl-cap: Sample of Cleaned Voting Locations Data
-
-# code referenced from: https://github.com/christina-wei/INF3104-1-Covid-Clinics/blob/main/outputs/paper/covid_clinics.qmd
-# & https://tellingstorieswithdata.com/20-r_essentials.html 
-head (cleaned_voting_locations, 5) |>
-  select(point_short_code, point_long_code, address_full) |>
-  kable(
-    col.names = c("Subdivision", "Ward & Subdivision", "Address"),
-  booktabs = TRUE
-  )
-
-```
-
 ### Election Voter Statistics
@@
 
-For the purpose of this paper, I did not include the Advance Vote and Mail In Voting subdivisions when counting the total number of subdivisions per ward.
+For the purpose of this paper, I included the Advance Vote and Mail In Voting subdivisions when counting the total number of subdivisions per ward.
 Firstly, there are only two Advance Vote locations distributed across every ward, meaning residents have to be motivated to travel farther away from their home to cast their ballot.

### 2023-06-02 scripts/02-data_cleaning.R
More edits to the script to summarize number of subdivisions

@@ -61,7 +61,7 @@ summarized_voter_statistics =
   ) |>
   arrange(desc(ward)) |>
   arrange(desc(n)) |>
-  select(ward, n)
+  select(ward, n) 
 summarized_voter_statistics
 
 # save summarized voter statistics data #

### 2023-06-02 scripts/02-data_cleaning.R
Updated script based on script I wrote in the data simulation section to validate the cleaned data

@@ -61,7 +61,8 @@ summarized_voter_statistics =
   ) |>
   arrange(desc(ward)) |>
   arrange(desc(n)) |>
-  select(ward, n) 
+  select(ward, n) |>
+  filter(!grepl('total', ward))
 summarized_voter_statistics
 
 # save summarized voter statistics data #

### 2023-06-02 scripts/02-data_cleaning.R
using similar code to the data simulation script, I updated the script to test the cleaned voter statistics data

@@ -61,8 +61,7 @@ summarized_voter_statistics =
   ) |>
   arrange(desc(ward)) |>
   arrange(desc(n)) |>
-  select(ward, n) |>
-  filter(!grepl('total', ward))
+  select(ward, n)
 summarized_voter_statistics
 
 # save summarized voter statistics data #
@@ -74,11 +73,18 @@ write_csv(
 
 #### Data validation ####
 
-# Check that there are between 1 and 25 wards #
-# Referenced code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
-summarized_voter_statistics$ward |> min() == 1
-summarized_voter_statistics$ward |> max() == 25
+# Check that there are no more than 25 wards #
+# Referenced code from: https://github.com/christina-wei/INF3104-1-Covid-Clinics/blob/main/scripts/00-simulation.R
+summarized_voter_statistics |>
+  group_by(ward) |>
+  count() |>
+  filter(n > ward) |>
+  sum() == 0
 
 # Check that number of subdivisions per ward is between 1 and 94
 summarized_voter_statistics$n |> min() == 1
 summarized_voter_statistics$n |> max() == 94
+
+# Check that voter turnout is between 11% and 100% #
+cleaned_voter_statistics$percent_voted |> min() == 1
+cleaned_voter_statistics$percent_voted|> max() == 100

### 2023-06-02 scripts/02-data_cleaning.R
Removed script that did not add the voter percentage accurately

@@ -59,11 +59,11 @@ summarized_voter_statistics =
   count(
     total = max(sub, na.rm = TRUE) 
   ) |>
-  arrange(desc(ward)) |>
+   arrange(desc(ward)) |>
   arrange(desc(n)) |>
   select(ward, n)
 summarized_voter_statistics
-
+    
 # save summarized voter statistics data #
 # based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
 write_csv(

### 2023-06-02 outputs/paper/paper.qmd
Updated information about the voter statistics data set and the process I undertook to clean the data set.

@@
 
+For the purpose of this paper, I included the Advance Vote and Mail In Voting subdivisions when counting the total number of subdivisions per ward.
+Firstly, there are only two Advance Vote locations distributed across an entire ward, meaning residents have to be motivated to travel farther away from their home to cast their ballot.
+Secondly, Mail In Voting requires applying online or by phone through Toronto Elections and then returning the completed Mail In Voting package by dropping it in a Canada Post mailbox or hand delivering it to one of the Toronto Elections drop boxes at select locations [@ElxnAccessibility].
+Voters need to be more organized and motivated to vote in advance of election day.
+
 Upon analysis, it was discovered that this data set included columns and data beyond the scope of this paper.
 I conducted basic cleaning of the data set to simplify column names and eliminate additional information (see @tbl-cleaned_voter_statistics).
-
-For the purpose of this paper, I included the Advance Vote and Mail In Voting subdivisions when counting the total number of subdivisions per ward.
-Firstly, there are only two Advance Vote locations distributed across every ward, meaning residents have to be motivated to travel farther away from their home to cast their ballot.
-Secondly, Mail In Voting requires applying online through Toronto Elections and then returning the completed Mail In Voting package by dropping it in a Canada Post mailbox or hand delivering it to one of the Toronto Elections drop boxes at select locations.
-Voters need to be more organized and motivated to vote in advance of election day.
 
@@
   )
+```
 
+I further cleaned the cleaned voter statistics data set to count the number of subdivision per ward and organize the number of subdivisions in descending order from the highest number of subdivisions to the lowest number of subdivisions.
+
+```{r}
+#| message: false
+#| echo: false
+
+head(summarized_voter_statistics, 25) |>
+  kable(
+    col.names = c("Ward", "Number of Subdivisions"), 
+    booktabs = TRUE
+  )
 ```

### 2023-06-02 scripts/02-data_cleaning.R
Further edits to the script summarizing the number of subdivisions per ward

@@ -55,15 +55,16 @@ write_csv(
 #Filter by ward and count number of subdivisions 
 summarized_voter_statistics = 
   cleaned_voter_statistics |>
- group_by(ward) |>
+  select(ward, sub, percent_voted) |>
+  group_by(ward) |>
   count(
     total = max(sub, na.rm = TRUE) 
   ) |>
-   arrange(desc(ward)) |>
-  arrange(desc(n)) |>
-  select(ward, n)
+  select(ward, n) |>
+  arrange(desc(ward)) |>
+  arrange(desc(n)) 
 summarized_voter_statistics
-    
+
 # save summarized voter statistics data #
 # based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
 write_csv(

### 2023-06-02 scripts/02-data_cleaning.R
Updated data validation tests

@@ -87,5 +87,5 @@ summarized_voter_statistics$n |> min() == 1
 summarized_voter_statistics$n |> max() == 94
 
 # Check that voter turnout is between 11% and 100% #
-cleaned_voter_statistics$percent_voted |> min() == 1
+cleaned_voter_statistics$percent_voted |> min() == 11
 cleaned_voter_statistics$percent_voted|> max() == 100

### 2023-06-02 outputs/paper/paper.qmd
Added in summarized voter statistics data

@@
 cleaned_voter_statistics <- read.csv(here::here("inputs/data/cleaned_voter_statistics.csv"))
+show_col_types = FALSE
+
+#read in cleaned data #
+summarized_voter_statistics <- read.csv(here::here("inputs/data/summarized_voter_statistics.csv"))
 show_col_types = FALSE

### 2023-06-02 scripts/02-data_cleaning.R
Updated script to filter out the total number lines for each ward

@@ -55,14 +55,13 @@ write_csv(
 #Filter by ward and count number of subdivisions 
 summarized_voter_statistics = 
   cleaned_voter_statistics |>
-  select(ward, sub, percent_voted) |>
+  select(ward, sub) |>
   group_by(ward) |>
   count(
     total = max(sub, na.rm = TRUE) 
   ) |>
   select(ward, n) |>
-  arrange(desc(ward)) |>
-  arrange(desc(n)) 
+  filter(n > 1) 
 summarized_voter_statistics
 
 # save summarized voter statistics data #

### 2023-06-02 scripts/02-data_cleaning.R
Added code to make ward numbers numeric which makes them appear in the correct order in the summarized data set.

@@ -39,6 +39,9 @@ cleaned_voter_statistics <-
     total_eligible_electors,
     number_voted,
     percent_voted
+  ) |>
+  mutate(
+    ward = as.numeric(ward)
   )
 head(cleaned_voter_statistics)
 

### 2023-06-02 scripts/02-data_cleaning.R
Change the number to be filtered out to 26 to remove the additional line at the bottom of the data set

@@ -64,7 +64,7 @@ summarized_voter_statistics =
     total = max(sub, na.rm = TRUE) 
   ) |>
   select(ward, n) |>
-  filter(n > 1) 
+  filter(n > 26) 
 summarized_voter_statistics
 
 # save summarized voter statistics data #
@@ -84,8 +84,8 @@ summarized_voter_statistics |>
   filter(n > ward) |>
   sum() == 0
 
-# Check that number of subdivisions per ward is between 1 and 94
-summarized_voter_statistics$n |> min() == 1
+# Check that number of subdivisions per ward is between 38 and 94
+summarized_voter_statistics$n |> min() == 38
 summarized_voter_statistics$n |> max() == 94
 
 # Check that voter turnout is between 11% and 100% #

### 2023-06-02 scripts/02-data_cleaning.R
Added script to add the summarized voter turnout by ward to the number of subdivisions summary

@@ -42,8 +42,14 @@ cleaned_voter_statistics <-
   ) |>
   mutate(
     ward = as.numeric(ward)
-  )
+  ) |>
 head(cleaned_voter_statistics)
+
+cleaned_voter_statistics |>
+  select(percent_voted) |>
+  mutate(
+  percent_voted = as.numeric(percent_voted)
+  )
 
 # save cleaned voter statistics data #
 # based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
@@ -66,6 +72,16 @@ summarized_voter_statistics =
   select(ward, n) |>
   filter(n > 26) 
 summarized_voter_statistics
+
+# Summarize voter turnout by ward
+summarized_voter_turnout = 
+cleaned_voter_statistics |>
+  select(ward, percent_voted) |>
+  group_by(ward) |>
+    count(
+      turnout = mean(percent_voted, na.rm = TRUE) 
+    ) 
+  summarized_voter_turnout
 
 # save summarized voter statistics data #
 # based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html

### 2023-06-02 outputs/paper/paper.qmd
changes to graph

@@
 
-head(summarized_voter_statistics, 25) |>
+summarized_voter_statistics |>
   kable(
@@
 #code referenced from: https://github.com/christina-wei/INF3104-1-Covid-Clinics/blob/main/outputs/paper/covid_clinics.qmd
-cleaned_voter_statistics |>
-  ggplot(aes(x = ward), "-", sub, fill = location_type) + 
+summarized_voter_statistics |>
+  ggplot(aes(x = ward), "-", sub, fill = n) + 
   geom_bar() + #bar graph

### 2023-06-02 scripts/02-data_cleaning.R
Continuing to edit the script to summarize voter turnout by ward

@@ -48,7 +48,7 @@ head(cleaned_voter_statistics)
 cleaned_voter_statistics |>
   select(percent_voted) |>
   mutate(
-  percent_voted = as.numeric(percent_voted)
+  percent_voted = as.integer(percent_voted)
   )
 
 # save cleaned voter statistics data #
@@ -79,7 +79,7 @@ cleaned_voter_statistics |>
   select(ward, percent_voted) |>
   group_by(ward) |>
     count(
-      turnout = mean(percent_voted, na.rm = TRUE) 
+      turnout = mean(percent_voted, na.rm = TRUE)
     ) 
   summarized_voter_turnout
 

### 2023-06-02 scripts/02-data_cleaning.R
Continuing to write code to summarize voter turnout by ward

@@ -48,7 +48,7 @@ head(cleaned_voter_statistics)
 cleaned_voter_statistics |>
   select(percent_voted) |>
   mutate(
-  percent_voted = as.integer(percent_voted)
+    percent_voted = as.numeric(percent_voted)
   )
 
 # save cleaned voter statistics data #
@@ -81,6 +81,14 @@ cleaned_voter_statistics |>
     count(
       turnout = mean(percent_voted, na.rm = TRUE)
     ) 
+  summarized_voter_turnout
+  
+  summarized_voter_turnout = 
+    cleaned_voter_statistics |>
+    select(ward, percent_voted) |>
+    group_by(ward) |>
+    count(
+    percent_voted)
   summarized_voter_turnout
 
 # save summarized voter statistics data #

### 2023-06-02 outputs/paper/paper.qmd
Updating code to graph results

@@
 
-#code referenced from: https://github.com/christina-wei/INF3104-1-Covid-Clinics/blob/main/outputs/paper/covid_clinics.qmd
 summarized_voter_statistics |>
-  ggplot(aes(x = ward), "-", sub, fill = n) + 
-  geom_bar() + #bar graph
-  theme_minimal() + #neater theme
-  theme(axis.text.x = element_text(angle=90, size = 6)) + #rotate x-axis & smaller font
+  ggplot(mapping = aes(x = ward, y = n)) +
+  geom_bar(stat = "identity") +
   labs(
-    title = "Number of Subdivisions Per Ward in Toronto",
+    title = "Number of Subdivisions per Ward",
     x = "Ward",
     y = "Number of Subdivisions",
-    fill = "Type"
-  )  #a
+  ) +
+  theme_classic() +
+  scale_fill_brewer(palette = "Set1")
 
@@
 
-head (cleaned_voter_statistics, 60) |>
- select(ward, sub) |>
-  arrange (ward) |>
+head (summarized_voter_statistics, 60) |>
    kable(

### 2023-06-02 scripts/02-data_cleaning.R
Wrote code to change the percent voted from character vector to numeric

@@ -45,11 +45,14 @@ cleaned_voter_statistics <-
   ) |>
 head(cleaned_voter_statistics)
 
+# Convert percent voted from character vector to numeric
+# Code referenced from: https://stackoverflow.com/questions/48273085/how-to-convert-a-character-vector-to-numeric
 cleaned_voter_statistics |>
   select(percent_voted) |>
   mutate(
-    percent_voted = as.numeric(percent_voted)
-  )
+    percent_voted = as.numeric(factor(percent_voted))
+)
+  
 
 # save cleaned voter statistics data #
 # based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
@@ -71,7 +74,10 @@ summarized_voter_statistics =
   ) |>
   select(ward, n) |>
   filter(n > 26) 
-summarized_voter_statistics
+summarized_voter_statistics |>
+rename(
+ subdivisions = n,
+) 
 
 # Summarize voter turnout by ward
 summarized_voter_turnout = 
@@ -88,7 +94,8 @@ cleaned_voter_statistics |>
     select(ward, percent_voted) |>
     group_by(ward) |>
     count(
-    percent_voted)
+      Turnout = mean(percent_voted)
+    )
   summarized_voter_turnout
 
 # save summarized voter statistics data #

### 2023-06-03 scripts/02-data_cleaning.R
still adjusting the code to summarize voter turnout by ward

@@ -53,7 +53,6 @@ cleaned_voter_statistics |>
     percent_voted = as.numeric(factor(percent_voted))
 )
   
-
 # save cleaned voter statistics data #
 # based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
 write_csv(
@@ -80,16 +79,7 @@ rename(
 ) 
 
 # Summarize voter turnout by ward
-summarized_voter_turnout = 
-cleaned_voter_statistics |>
-  select(ward, percent_voted) |>
-  group_by(ward) |>
-    count(
-      turnout = mean(percent_voted, na.rm = TRUE)
-    ) 
-  summarized_voter_turnout
-  
-  summarized_voter_turnout = 
+ summarized_voter_turnout = 
     cleaned_voter_statistics |>
     select(ward, percent_voted) |>
     group_by(ward) |>

### 2023-06-03 scripts/02-data_cleaning.R
changes to summarizing voter turnout by ward code

@@ -79,7 +79,7 @@ rename(
 ) 
 
 # Summarize voter turnout by ward
- summarized_voter_turnout = 
+summarized_voter_turnout = 
     cleaned_voter_statistics |>
     select(ward, percent_voted) |>
     group_by(ward) |>

### 2023-06-03 scripts/02-data_cleaning.R
Further edits to the code trying to summarize voter turnout by ward

@@ -80,13 +80,14 @@ rename(
 
 # Summarize voter turnout by ward
 summarized_voter_turnout = 
-    cleaned_voter_statistics |>
-    select(ward, percent_voted) |>
-    group_by(ward) |>
-    count(
-      Turnout = mean(percent_voted)
-    )
-  summarized_voter_turnout
+  cleaned_voter_statistics |>
+  select(ward, percent_voted) |>
+  group_by(ward) |>
+  count(
+    Turnout = mean(percent_voted)
+  )
+summarized_voter_turnout
+
 
 # save summarized voter statistics data #
 # based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html

### 2023-06-03 outputs/paper/paper.qmd
Updated code to graph number of subdivisions per ward

@@
 
-summarized_voter_statistics |>
+head(summarized_voter_statistics, 5) |>
   kable(
@@
 #| echo: false
+#| label: fig-voter_statistics
+#| tbl-cap: Number of Subdivisions per Ward
 
+#Code refernced from: https://stackoverflow.com/questions/11335836/increase-number-of-axis-ticks
 summarized_voter_statistics |>
@@
   geom_bar(stat = "identity") +
+scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
+scale_y_continuous(breaks = scales::pretty_breaks(n = 25))+
   labs(
@@
   ) +
-  theme_classic() +
-  scale_fill_brewer(palette = "Set1")
+  theme_classic()
 
@@
 
-cleaned_voter_statistics |>
-  slice(2:52) |>
-  count(
-    total = max(sub, na.rm = TRUE)
-  )
-
 ```

### 2023-06-03 outputs/paper/paper.qmd
Added to the introduction, data, and results sections

@@
 Coming out of the height of the Covid-19 pandemic, the 2022 election largely upheld the status quo and did not feature ambitious policy platforms.
-Voter turnout across the city was 29.7% - the lowest in the city’s history since amalgamation in 1997, with turnout ranging on a ward-by-ward basis from 22.1% to 38.5% [@Turnout22; @DismalTurnout].
+Voter turnout across the city was 29.7% - the lowest in the city’s history since amalgamation in 1997, with turnout ranging on a ward-by-ward basis from 22% to 38% [@Turnout22; @DismalTurnout].
 
@@
 The data utilized throughout this paper was obtained through the City of Toronto’s OpenDataToronto Library [@rOpenDataToronto].
-The two different data sets used are: “Elections Voting Locations” [@VotingLocations] and “Elections - Voter Statistics” [@VoterStatistics].
+The data sets used is entitled “Elections - Voter Statistics” [@VoterStatistics].
 Data was collected and analyzed using the statistical programming software R (@r), with additional support from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], and `here` [@rHere].
@@
 
-I further cleaned the cleaned voter statistics data set to count the number of subdivision per ward and organize the number of subdivisions in descending order from the highest number of subdivisions to the lowest number of subdivisions.
+Moreover, I further cleaned the voter statistics data set to count the total number of subdivisions per ward and display the results next to the ward number (see @tbl-summarized_voter_statistics).
 
@@
 #| echo: false
+#| #| label: tbl-summarized_voter_statistics
+#| tbl-cap: Sample of Summarized Voter Statistics Data
 
@@
 ## Results
+During the 2022 municipal election, there were 1,535 subdivisions across Toronto's 25 wards.
+Ward 10 had the highest number of subdivisions at 94, while Ward 23 had the lowest number at 38.
+Ward 13 had the second highest number of subdivisions at 90, with Ward 3 coming in third highest with 83 subdivisions.
+Ward 7 had the second lowest number of subdivisions at 47, following by Ward 25 with 48.
+The range of subdivisions within Wards varied more than anticipated, given that ward boundaries and their corresponding population are relatively equal.
 

### 2023-06-03 outputs/paper/paper.qmd
Updated results section with ward names

@@
 During the 2022 municipal election, there were 1,535 subdivisions across Toronto's 25 wards.
-Ward 10 had the highest number of subdivisions at 94, while Ward 23 had the lowest number at 38.
-Ward 13 had the second highest number of subdivisions at 90, with Ward 3 coming in third highest with 83 subdivisions.
-Ward 7 had the second lowest number of subdivisions at 47, following by Ward 25 with 48.
+Ward 10 (Spadina-Fort York) had the highest number of subdivisions at 94, while Ward 23 (Scarborough North) had the lowest number at 38.
+Ward 13 (Toronto Centre) had the second highest number of subdivisions at 90, with Ward 3 (Etobicoke-Lakeshore) coming in third highest with 83 subdivisions.
+Ward 7 (Humber River-Black Creek) had the second lowest number of subdivisions at 47, following by Ward 25 (Scarborough-Rouge Park) with 48.
 The range of subdivisions within Wards varied more than anticipated, given that ward boundaries and their corresponding population are relatively equal.

### 2023-06-03 outputs/paper/paper.qmd
Updated script to fix numbering of x axis on the number of subdivisions per ward graph

@@
   geom_bar(stat = "identity") +
-scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
+  scale_x_continuous(breaks = scales::pretty_breaks(n = 25)) +
   scale_y_continuous(breaks = scales::pretty_breaks(n = 25))+

### 2023-06-03 scripts/02-data_cleaning.R
Editing script to summarize voter turnout by ward

@@ -84,7 +84,7 @@ summarized_voter_turnout =
   select(ward, percent_voted) |>
   group_by(ward) |>
   count(
-    Turnout = mean(percent_voted)
+    Turnout = mean(percent_voted) 
   )
 summarized_voter_turnout
 

### 2023-06-03 scripts/02-data_cleaning.R
Continuing to update script to summarize voter turnout by ward

@@ -79,16 +79,18 @@ rename(
 ) 
 
 # Summarize voter turnout by ward
-summarized_voter_turnout = 
-  cleaned_voter_statistics |>
+summarized_voter_statistics =
+cleaned_voter_statistics |>
   select(ward, percent_voted) |>
+  mutate(
+    percent_voted = as.numeric(factor(percent_voted))
+  ) |>
   group_by(ward) |>
   count(
-    Turnout = mean(percent_voted) 
+    Turnout = mean(percent_voted, na.rm = TRUE) 
   )
-summarized_voter_turnout
-
-
+summarized_voter_statistics
+  
 # save summarized voter statistics data #
 # based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
 write_csv(

### 2023-06-03 outputs/paper/paper.qmd
Edits to introduction, data, and election voter statistics sections

@@
 library(ggplot2)
-library(lubridate)
 
@@
 To examine the impact of subdivision placement on voter turnout, this paper is organized into the following sections: Data, Results, Discussion, and Conclusion.
-In the Data section, I discuss the nature of the spreadsheets obtained through the City of Toronto’s OpenDataToronto Library [@rOpenDataToronto] and the steps I took to clean and analyze the data.
+In the Data section, I discuss the nature of the spreadsheet obtained through the City of Toronto’s OpenDataToronto Library [@rOpenDataToronto] and the steps I took to clean and analyze the data.
 The Results section highlights trends found during the analysis process, while the Discussion section further evaluates the trends and presents insight.
@@
 The data utilized throughout this paper was obtained through the City of Toronto’s OpenDataToronto Library [@rOpenDataToronto].
-The data sets used is entitled “Elections - Voter Statistics” [@VoterStatistics].
-Data was collected and analyzed using the statistical programming software R (@r), with additional support from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], and `here` [@rHere].
+The data set used is entitled “Elections - Voter Statistics” [@VoterStatistics].
+Data was collected and analyzed using the statistical programming software R (@r), with additional support from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], `KableExtra` [@rKableExtra], `Knitr` [@rknitr], and `here` [@rHere].
 A further discussion of the data collection, cleaning, and analysis process can be found later on in this paper.
@@
 This dataset, published by the City Clerk's Office [@VoterStatistics] comprises the number of eligible voters and corrections/additions to the Voter’s List for the 2022 municipal election.
-The dataset is broken down by ward and voting subdivision, including voting for school board trustees.
+The dataset is broken down by ward and voting subdivision.
 This data set was last refreshed on February 7, 2023 and captured for this paper on May 21, 2023.
-This data set outlines the total number of eligible electors in Toronto, by ward, subdivision, and the entire city.
+The total number of eligible electors in Toronto, by ward, subdivision, and the entire city are outlined in this data set.
 

### 2023-06-03 scripts/02-data_cleaning.R
Futher edits to the script to summarize voter turnout and data validation

@@ -79,17 +79,17 @@ rename(
 ) 
 
 # Summarize voter turnout by ward
-summarized_voter_statistics =
+summarized_voter_turnout =
 cleaned_voter_statistics |>
   select(ward, percent_voted) |>
   mutate(
     percent_voted = as.numeric(factor(percent_voted))
   ) |>
   group_by(ward) |>
-  count(
-    Turnout = mean(percent_voted, na.rm = TRUE) 
+  summarise(
+    Turnout = mean(percent_voted, na.rm = TRUE)
   )
-summarized_voter_statistics
+summarized_voter_turnout
   
 # save summarized voter statistics data #
 # based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
@@ -109,8 +109,8 @@ summarized_voter_statistics |>
   sum() == 0
 
 # Check that number of subdivisions per ward is between 38 and 94
-summarized_voter_statistics$n |> min() == 38
-summarized_voter_statistics$n |> max() == 94
+summarized_voter_statistics$subdivisions |> min() == 38
+summarized_voter_statistics$subdivisions |> max() == 94
 
 # Check that voter turnout is between 11% and 100% #
 cleaned_voter_statistics$percent_voted |> min() == 11

### 2023-06-03 outputs/paper/paper.qmd
Edits to the code to display graphs

@@
 #| echo: false
+#| #| label: fig-voter_statistics
+#| tbl-cap: Number of Subdivisions per Ward
 
 # summarize ward data by number of subdivisions
-
-head (summarized_voter_statistics, 60) |>
+summarized_voter_statistics |>
    kable(
-    col.names = c("Ward", "Sub"),
+    col.names = c("Ward", "Number of Subdivisions"),
   booktabs = TRUE

### 2023-06-03 scripts/02-data_cleaning.R
Added code to extract the voter turnout by ward on a row by row basis

@@ -88,7 +88,20 @@ cleaned_voter_statistics |>
   group_by(ward) |>
   summarise(
     Turnout = mean(percent_voted, na.rm = TRUE)
-  )
+  ) 
+summarized_voter_turnout
+
+#Find voter turnout by ward
+summarized_voter_turnout = 
+  transpose((cleaned_voter_statistics[56,]))[[1]] 
+summarized_voter_turnout
+
+summarized_voter_turnout = 
+  transpose((cleaned_voter_statistics[129,]))[[1]] 
+summarized_voter_turnout
+
+summarized_voter_turnout = 
+  transpose((cleaned_voter_statistics[216,]))[[1]] 
 summarized_voter_turnout
   
 # save summarized voter statistics data #

### 2023-06-03 scripts/02-data_cleaning.R
Testing script to summarize voter turnout by ward

@@ -93,7 +93,7 @@ summarized_voter_turnout
 
 #Find voter turnout by ward
 summarized_voter_turnout = 
-  transpose((cleaned_voter_statistics[56,]))[[1]] 
+  transpose((cleaned_voter_statistics[56,]))[[1]]
 summarized_voter_turnout
 
 summarized_voter_turnout = 

### 2023-06-03 scripts/02-data_cleaning.R
Removed script I had been playing around with to summarize voter turnout by ward and added correct script. Also, updated data validation script to reflect the cleaned data set.

@@ -78,37 +78,26 @@ rename(
  subdivisions = n,
 ) 
 
-# Summarize voter turnout by ward
-summarized_voter_turnout =
-cleaned_voter_statistics |>
-  select(ward, percent_voted) |>
-  mutate(
-    percent_voted = as.numeric(factor(percent_voted))
-  ) |>
-  group_by(ward) |>
-  summarise(
-    Turnout = mean(percent_voted, na.rm = TRUE)
-  ) 
-summarized_voter_turnout
-
-#Find voter turnout by ward
-summarized_voter_turnout = 
-  transpose((cleaned_voter_statistics[56,]))[[1]]
-summarized_voter_turnout
-
-summarized_voter_turnout = 
-  transpose((cleaned_voter_statistics[129,]))[[1]] 
-summarized_voter_turnout
-
-summarized_voter_turnout = 
-  transpose((cleaned_voter_statistics[216,]))[[1]] 
-summarized_voter_turnout
-  
 # save summarized voter statistics data #
 # based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
 write_csv(
   x = summarized_voter_statistics,
   file = "summarized_voter_statistics.csv"
+)
+
+# Add voter turnout by ward from cleaned_voter_statistics data set
+summarized_voter_turnout <-
+  data <- data.frame(
+    ward=c("1","2","3","4","5", "6", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16", "17", "18", "19", "20", "21", "22", "23", "24", "25") ,  
+    turnout=c(24, 30, 33, 38, 28, 25, 22, 32, 31, 24, 32, 34, 29, 36, 33, 31, 27, 29, 36, 32, 26, 26, 25, 27, 27)
+  )
+summarized_voter_turnout
+  
+# save summarized voter turnout data #
+# based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
+write_csv(
+  x = summarized_voter_turnout,
+  file = "summarized_voter_turnout.csv"
 )
 
 #### Data validation ####
@@ -122,9 +111,9 @@ summarized_voter_statistics |>
   sum() == 0
 
 # Check that number of subdivisions per ward is between 38 and 94
-summarized_voter_statistics$subdivisions |> min() == 38
-summarized_voter_statistics$subdivisions |> max() == 94
+summarized_voter_statistics$n |> min() == 38
[…]

### 2023-06-03 scripts/02-data_cleaning.R
Added script to merge the two cleaned data sets which had number of subdivisions and turnout per ward

@@ -100,6 +100,22 @@ write_csv(
   file = "summarized_voter_turnout.csv"
 )
 
+# Merge number of subdivisions and voter turnout
+finalized_voter_statistics = 
+  merge(
+    summarized_voter_turnout,
+    summarized_voter_statistics,
+    by = "ward"
+  )
+finalized_voter_statistics
+
+# save finalized voter statistics data #
+# based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
+write_csv(
+  x = finalized_voter_statistics,
+  file = "finalized_voter_statistics.csv"
+)
+
 #### Data validation ####
 
 # Check that there are no more than 25 wards #

### 2023-06-03 scripts/02-data_cleaning.R
Updated data validation tests to reflect merged data sets

@@ -120,16 +120,16 @@ write_csv(
 
 # Check that there are no more than 25 wards #
 # Referenced code from: https://github.com/christina-wei/INF3104-1-Covid-Clinics/blob/main/scripts/00-simulation.R
-summarized_voter_statistics |>
+finalized_voter_statistics |>
   group_by(ward) |>
   count() |>
   filter(n > ward) |>
   sum() == 0
 
 # Check that number of subdivisions per ward is between 38 and 94
-summarized_voter_statistics$n |> min() == 38
-summarized_voter_statistics$n |> max() == 94
+finalized_voter_statistics$n |> min() == 38
+finalized_voter_statistics$n |> max() == 94
 
 # Check that voter turnout is between 22% and 38% #
-summarized_voter_turnout$turnout |> min() == 22
-summarized_voter_turnout$turnout|> max() == 38
+finalized_voter_statistics$turnout |> min() == 22
+finalized_voter_statistics$turnout|> max() == 38

### 2023-06-03 outputs/paper/paper.qmd
Formatting of data section by adding subheaders and editing the text for each section

--- 2023-06-03 outputs/paper/paper.qmd
@@
 ---
-title: "Do More Subdivisions in Wards Increase Turnout? An Analysis of the 2022 Toronto Municipal Election"
+title: "Do More Subdivisions in Wards Increase Voter Turnout? An Analysis of the 2022 Toronto Municipal Election"
 author: "Inessa De Angelis"
@@
 show_col_types = FALSE
+
+#read in cleaned data #
+summarized_voter_turnout <- read.csv(here::here("inputs/data/summarized_voter_turnout.csv"))
+show_col_types = FALSE
+
+#read in cleaned data #
+finalized_voter_statistics <- read.csv(here::here("inputs/data/finalized_voter_statistics.csv"))
+show_col_types = FALSE
 ```
@@
 As mandated by the Government of Ontario, residents of the City of Toronto went to the polls on October 24, 2022 to elect a mayor, councillors, and school board trustees.
-Then-sitting Mayor, John Tory sought re-election, along with a number of incumbent city councillors.
+Then-sitting Mayor, John Tory, sought re-election, along with a number of incumbent city councillors.
 Coming out of the height of the Covid-19 pandemic, the 2022 election largely upheld the status quo and did not feature ambitious policy platforms.
-Voter turnout across the city was 29.7% - the lowest in the city’s history since amalgamation in 1997, with turnout ranging on a ward-by-ward basis from 22% to 38% [@Turnout22; @DismalTurnout].
+Voter turnout across the city fell to 30% - the lowest in the city’s history since amalgamation in 1997, with turnout ranging on a ward-by-ward basis from 22% to 38% [@Turnout22; @DismalTurnout].
 
-Individual turnout by ward largely depended on local candidates and the accessibility of subdivisions to cast a ballot.
+Individual turnout by ward depended on local candidates and the accessibility of subdivisions to cast a ballot.
 The City of Toronto @ElectionDictionary defines a ward as “a geographical area represented by a member of Council.” Following Premier Doug Ford’s decision in the summer of 2018 to cut down the number of wards in the city to match federal and provincial riding boundaries, there are currently 25 wards in the City of Toronto [@BigCityElxns].
@@
 The data utilized throughout this paper was obtained through the City of Toronto’s OpenDataToronto Library [@rOpenDataToronto].
-The data set used is entitled “Elections - Voter Statistics” [@VoterStatistics].
+The data set used is entitled 'Elections - Voter Statistics' [@VoterStatistics].
 Data was collected and analyzed using the statistical programming software R (@r), with additional support from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], `KableExtra` [@rKableExtra], `Knitr` [@rknitr], and `here` [@rHere].
@@
 ### Election Voter Statistics
-This dataset, published by the City Clerk's Office [@VoterStatistics] comprises the number of eligible voters and corrections/additions to the Voter’s List for the 2022 municipal election.
-The dataset is broken down by ward and voting subdivision.
-This data set was last refreshed on February 7, 2023 and captured for this paper on May 21, 2023.
-The total number of eligible electors in Toronto, by ward, subdivision, and the entire city are outlined in this data set.
+This dataset, published by the City Clerk's Office [@VoterStatistics] outlines voter statistics on a ward-by-ward and entire city basis for the 2022 municipal election.
+For each subdivision within a ward, the data set shows the polling location name and address, number of additions and corrections to the voter's list, number of eligible electors and number who voted, and rejected and declined ballots.This data set was last refreshed on February 7, 2023 and captured for this paper on May 21, 2023.
 
-For the purpose of this paper, I included the Advance Vote and Mail In Voting subdivisions when counting the total number of subdivisions per ward.
-Firstly, there are only two Advance Vote locations distributed across an entire ward, meaning residents have to be motivated to travel farther away from their home to cast their ballot.
-Secondly, Mail In Voting requires applying online or by phone through Toronto Elections and then returning the completed Mail In Voting package by dropping it in a Canada Post mailbox or hand delivering it to one of the Toronto Elections drop boxes at select locations [@ElxnAccessibility].
-Voters need to be more organized and motivated to vote in advance of election day.
-
-Upon analysis, it was discovered that this data set included columns and data beyond the scope of this paper.
-I conducted basic cleaning of the data set to simplify column names and eliminate additional information (see @tbl-cleaned_voter_statistics).
+Upon analysis, columns such as rejected and declined ballots were deemed beyond the scope of this paper.
+I conducted the first step of basic data cleaning to eliminate additional columns and simplify the names of retained columns (see @tbl-cleaned_voter_statistics).
 
@@
 ```
+### Number of Subdivisions 
+For the purpose of this paper, I included the Advance Vote and Mail In Voting subdivisions when counting the total number of subdivisions per ward.
+Firstly, there are only two Advance Vote locations distributed across an entire ward, meaning residents have to be motivated to travel farther away from their home to cast their ballot.
+Secondly, Mail In Voting requires applying online or by phone through Toronto Elections and then returning the completed Mail In Voting package by dropping it in a Canada Post mailbox or hand delivering it to one of the select Toronto Elections drop boxes [@ElxnAccessibility].
+Voters need to be organized and motivated to vote in advance of election day.
 
-Moreover, I further cleaned the voter statistics data set to count the total number of subdivisions per ward and display the results next to the ward number (see @tbl-summarized_voter_statistics).
+I further cleaned the voter statistics data set to count the total number of subdivisions per ward and display the results next to the ward number (see @tbl-summarized_voter_statistics).
 
@@
 ```
+
+### Turnout by Ward
+Moreover, using the cleaned voter statistics data set, I created a table with
+
+### Merged Number of Subdivisions and Turnout
 
@@
 #Code refernced from: https://stackoverflow.com/questions/11335836/increase-number-of-axis-ticks
-summarized_voter_statistics |>
+finalized_voter_statistics |>
  ggplot(mapping = aes(x = ward, y = n)) +
@@
 #| echo: false
+#| #| label: fig-turnout_by_ward
+#| tbl-cap: Voter Turnout by Ward
 
-# summarize ward data by number of subdivisions
-
-head (summarized_voter_statistics, 60) |>
-   kable(
-    col.names = c("Ward", "Sub"),
-  booktabs = TRUE
-  )
+#Voter turnout by ward
+#Code refernced from: https://stackoverflow.com/questions/11335836/increase-number-of-axis-ticks
+finalized_voter_statistics |>
+ ggplot(mapping = aes(x = ward, y = turnout)) +
+  geom_bar(stat = "identity") +
+  scale_x_continuous(breaks = scales::pretty_breaks(n = 25)) +
+  scale_y_continuous(breaks = scales::pretty_breaks(n = 25))+
+  labs(
+    title = "Voter Turnout by Ward",
+    x = "Ward",
+    y = "Turnout (%)",
+  ) +
+  theme_classic()
 

### 2023-06-03 outputs/paper/paper.qmd
Continued editing the data sections and adding to the result sections. Also added the conclusion

@@
 The data set used is entitled 'Elections - Voter Statistics' [@VoterStatistics].
-Data was collected and analyzed using the statistical programming software R (@r), with additional support from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], `KableExtra` [@rKableExtra], `Knitr` [@rknitr], and `here` [@rHere].
+Data was collected and analyzed using the statistical programming software R [@r], with additional support from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], `KableExtra` [@rKableExtra], `Knitr` [@rknitr], and `here` [@rHere].
 A further discussion of the data collection, cleaning, and analysis process can be found later on in this paper.
@@
 This dataset, published by the City Clerk's Office [@VoterStatistics] outlines voter statistics on a ward-by-ward and entire city basis for the 2022 municipal election.
-For each subdivision within a ward, the data set shows the polling location name and address, number of additions and corrections to the voter's list, number of eligible electors and number who voted, and rejected and declined ballots.This data set was last refreshed on February 7, 2023 and captured for this paper on May 21, 2023.
+For each subdivision within a ward, the data set shows the polling location name and address, number of additions and corrections to the voter's list, number of eligible electors and number who voted, and rejected and declined ballots.
+This data set was last refreshed on February 7, 2023 and captured for this paper on May 21, 2023.
 
@@
 ### Number of Subdivisions
-For the purpose of this paper, I included the Advance Vote and Mail In Voting subdivisions when counting the total number of subdivisions per ward.
+
+The Election Voter Statistics data set [@VoterStatistics] listed "NA" for the number of subdivisions per ward, but included the voter turnout average for all subdivisions within a ward.
+Consequently, I wrote a script using the `count` function to count the number of subdivisions per ward.
+I included the Advance Vote and Mail In Voting subdivisions when counting the total number of subdivisions per ward (see @tbl-summarized_voter_statistics).
 Firstly, there are only two Advance Vote locations distributed across an entire ward, meaning residents have to be motivated to travel farther away from their home to cast their ballot.
 Secondly, Mail In Voting requires applying online or by phone through Toronto Elections and then returning the completed Mail In Voting package by dropping it in a Canada Post mailbox or hand delivering it to one of the select Toronto Elections drop boxes [@ElxnAccessibility].
-Voters need to be organized and motivated to vote in advance of election day.
-
-I further cleaned the voter statistics data set to count the total number of subdivisions per ward and display the results next to the ward number (see @tbl-summarized_voter_statistics).
+Despite the number of subdivisions per ward, the City of Toronto has earmarked subdivision 97 for Mail In Voting, while subdivisions 98 and 99 are designated Advance Vote locations.
 
@@
 #| echo: false
-#| #| label: tbl-summarized_voter_statistics
+#| label: tbl-summarized_voter_statistics
 #| tbl-cap: Sample of Summarized Voter Statistics Data
@@
 ### Turnout by Ward
-Moreover, using the cleaned voter statistics data set, I created a table with
+Moreover, using the cleaned voter statistics data set, I added the voter turnout by ward percentage from the bottom of each ward section to a table I created with each ward number.
+To view a sample of this data set, please see @tbl-summarized_voter_turnout.
+
+```{r}
+#| message: false
+#| echo: false
+#| label: tbl-summarized_voter_turnout
+#| tbl-cap: Sample of Summarized Voter Turnout Data
+
+head(summarized_voter_turnout, 5) |>
+  kable(
+    col.names = c("Ward", "Turnout (%)"), 
+    booktabs = TRUE
+  )
+```
 
 ### Merged Number of Subdivisions and Turnout 
+Lastly, using the `merge` function, I created a script to combine the summarized number of subdivisions and voter turnout data into one data set.
+See @tbl-finalized_voter_turnout for a sample of this merged data set.
+
+```{r}
+#| message: false
+#| echo: false
+#| label: tbl-finalized_voter_turnout
+#| tbl-cap: Sample of Merged Subdivisions and Voter Turnout Data Set
+
+head(finalized_voter_statistics, 5) |>
+  kable(
+    col.names = c("Ward", "Turnout (%)", "Subdivisions"), 
+    booktabs = TRUE
+  )
+```
 
 ## Results
+### Subdivisions per Ward
 During the 2022 municipal election, there were 1,535 subdivisions across Toronto's 25 wards.
@@
 Ward 7 (Humber River-Black Creek) had the second lowest number of subdivisions at 47, following by Ward 25 (Scarborough-Rouge Park) with 48.
-The range of subdivisions within Wards varied more than anticipated, given that ward boundaries and their corresponding population are relatively equal.
+@fig-voter-statistics displays the total number of subdivisions for all 25 wards.
 
@@
 #| label: fig-voter_statistics
-#| tbl-cap: Number of Subdivisions per Ward
+#| fig-cap: Number of Subdivisions per Ward
 
@@
   theme_classic() 
+```
 
-```
+### Voter Turnout per Ward
+Voter turnout across the entire city during the 2022 municipal election fell to 30%, with Ward 4 (Parkdale-High Park) having the highest turnout at 38% and Ward 7 (Humber River-Black Creek) having the lowest at 22%.
+Ward 19 (Beaches-East York) and Ward 14 (Toronto Danforth) were tied for second highest turnout at 36% each.
+The third highest turnout at 34% was in Ward 12 (Toronto-St.
+Paul's).
+The wards with the second lowest turnout were Ward 1 (Etobicoke North) and Ward 10 (Spadina-Fork York) at 24% each.
+Ward 6 (York Centre) and Ward 23 (Scarborough North) with 25% turnout respectively were tied for third lowest turnout.
+For a full picture of voter turnout per ward, please refer to @fig-turnout_by_ward.
 
@@
 #| echo: false
-#| #| label: fig-turnout_by_ward
-#| tbl-cap: Voter Turnout by Ward
+#| label: fig-turnout_by_ward
+#| fig-cap: Voter Turnout by Ward
 
@@
  ggplot(mapping = aes(x = ward, y = turnout)) +
-  geom_bar(stat = "identity") +
   scale_x_continuous(breaks = scales::pretty_breaks(n = 25)) +
   scale_y_continuous(breaks = scales::pretty_breaks(n = 25))+
+  geom_bar(stat = "identity") +
   labs(
@@
   theme_classic() 
+```
+### Comparing Number of Subdivisions and Turnout
 
+```{r}
+#| message: false
+#| echo: false
+#| label: fig-comparing-sub-turnout
+#| fig-cap: Correlation between Number of Subdivisions and Voter Turnout
+
+#Voter turnout by ward
+ggplot(data = finalized_voter_statistics, mapping = aes(x = turnout, y = n)) +
+  geom_point(shape = 21, fill = '#0f993d', color = 'white', size = 3) +
+  sm_statCorr() 
 ```
+```
+
 ## Discussion
@@
 ## Conclusion
+This paper investigates the number of subdivisions within Toronto’s 25 wards in relation to voter turnout to see if the wards with more subdivisions had higher turnout during the 2022 municipal election.
+My results reveal that wards with more subdivisions did not correlate to higher voter turnout.
+Factors such as income, level of education attained, and an analysis of local races for councillor and school board trustee within a ward (whether there was an incumbent or not) would aid in understanding voter behaviour in municipal elections.
+Data from the upcoming Toronto Mayoral by-election on June 26, 2023 may reveal further insights into the placement of polling locations in relation to voter turnout.
+Cities and municipalities across Ontario are facing declining voter turnout and the factors must be identified and addressed, whether it be more strategically choosing subdivisions and polling locations in wards, introducing online voting mechanisms, or implementing democratic reform [@reducingcost].
+Strong local representation is crucial for the future of democracy in Toronto and subdivisions play an important role in ensuring voters have their voices heard.
 

### 2023-06-03 outputs/paper/paper.qmd
Formatting

@@
 ```
-```
 

### 2023-06-03 outputs/paper/paper.qmd
Added ggbeeswarm package, graph visualizing voter turnout in comparison to number of subdivisions, and added text to the discussion section

@@
 library(ggplot2)
+library(ggbeeswarm)
 
@@
 #Voter turnout by ward
-ggplot(data = finalized_voter_statistics, mapping = aes(x = turnout, y = n)) +
-  geom_point(shape = 21, fill = '#0f993d', color = 'white', size = 3) +
-  sm_statCorr() 
+#Code referenced from: https://github.com/christina-wei/INF3104-1-Covid-Clinics/blob/main/outputs/paper/covid_clinics.pdf
+finalized_voter_statistics |>
+  ggplot(aes(x = turnout, y = n, label = ward)) + 
+  geom_point() + 
+  geom_beeswarm() +
+  geom_smooth(method=lm, se=FALSE) + 
+  geom_text(hjust=-0.5, vjust=0.5) + 
+  labs(
+    title = "Number of Subdivisions vs Voter Turnout by Ward",
+    x = "Turnout (%)",
+    y = "Subdivisions",
+    label = "Ward"
+  ) +
+  theme_minimal()
 ```
@@
 ## Discussion
+Overall, the data reveals that having more subdivisions within wards does not automatically correlate to higher voter turnout, as illustrated by @fig-comparing-sub-turnout.
+Ward 4 (Parkdale-High Park) which had the highest voter turnout at 38% has 70 subdivisions within the ward, which falls approximately in the middle of the number of subdivisions across all 25 wards.
+Ward 7 (Humber River-Black Creek) had a 22% voter turnout and 47 subdivisions, emphasizing that having fewer subdivisions potentially contributed to having the lowest voter turnout across the city.
+Ward 10 (Spadina-Fork York), which had the second lowest voter turnout at 24%, accounts for the highest number of subdivisions within a ward at 94.
+Previous research reveals that younger, more transient communities reside in neighbourhoods in the Spadina-Fork York riding and are less likely to engage with municipal politics, potentially contributing to the low voter turnout despite the high number of subdivisions [@Turnout22].
+
+Furthermore, Ward 19 (Beaches-East York) and Ward 14 (Toronto Danforth) which both saw the second highest voter turnout at 36% had 57 and 52 subdivisions respectively, which is on the lower end of number of subdivisions.
 
@@
 My results reveal that wards with more subdivisions did not correlate to higher voter turnout.
-Factors such as income, level of education attained, and an analysis of local races for councillor and school board trustee within a ward (whether there was an incumbent or not) would aid in understanding voter behaviour in municipal elections.
+Factors such as income, level of education attained, and an analysis of local races for councillor and school board trustee within a ward (whether there was an incumbent or not) should be analyzed in further studies of voter behaviour during municipal elections.
 Data from the upcoming Toronto Mayoral by-election on June 26, 2023 may reveal further insights into the placement of polling locations in relation to voter turnout.

### 2023-06-03 outputs/paper/paper.qmd
Adding more to the discussion section

@@
 Ward 4 (Parkdale-High Park) which had the highest voter turnout at 38% has 70 subdivisions within the ward, which falls approximately in the middle of the number of subdivisions across all 25 wards.
-Ward 7 (Humber River-Black Creek) had a 22% voter turnout and 47 subdivisions, emphasizing that having fewer subdivisions potentially contributed to having the lowest voter turnout across the city.
+Ward 19 (Beaches-East York) and Ward 14 (Toronto Danforth) which both saw the second highest voter turnout at 36% had 57 and 52 subdivisions respectively, once again highlighting that more subdivisions does not necessarily correlate to higher voter turnout.
+
+Futhermore, Ward 7 (Humber River-Black Creek) had a 22% voter turnout and 47 subdivisions, emphasizing that having fewer subdivisions potentially contributed to having the lowest voter turnout across the city.
 Ward 10 (Spadina-Fork York), which had the second lowest voter turnout at 24%, accounts for the highest number of subdivisions within a ward at 94.
-Previous research reveals that younger, more transient communities reside in neighbourhoods in the Spadina-Fork York riding and are less likely to engage with municipal politics, potentially contributing to the low voter turnout despite the high number of subdivisions [@Turnout22].
+Previous research reveals that younger, more transient communities reside in the Spadina-Fork York riding and are less likely to engage with municipal politics, potentially contributing to the low voter turnout despite the high number of subdivisions [@Turnout22].
 
-Furthermore, Ward 19 (Beaches-East York) and Ward 14 (Toronto Danforth) which both saw the second highest voter turnout at 36% had 57 and 52 subdivisions respectively, which is on the lower end of number of subdivisions.
+The Wards within Scarborough (Wards 20-25) consistently had some of the lowest voter turnouts and fewest number of subdivisions across all 25 wards.
+Ward 23 (Scarborough North) witnessed the third lowest voter turnout across the entire city at 25% and the fewest number of subdivisions at 38.
+Ward 20 (Scarborough Southwest) had the highest turnout of all Scarborough wards at 32% and had 61 subdivisions.
+Scarborough's challenges with accessible public transit are well recognized, potentially contributing to the lower voter turnout.
+
+The creation of subdivision boundaries within wards lacks transparency and clarity.
+The City Clerk is responsible for drawing up all ward subdivisions.
+We don’t know what data and information they use to make these decisions and what types of bias may be introduced.
+Don’t know if there’s a public input process or if like creating ward boundaries, it is an individual who has little local knowledge of the very community impacted by the creation of boundaries
+
+Toronto Elections [@ElxnAccessibility] clearly outlines their process for finding and establishing accessible polling locations within subdivisions but does not elaborate on the process of establishing subdivisions themselves.
+
+There is a lack of clarity whether the people who voted at Advanced Polls or by Mail In Ballot were removed from the total number of electors by subdivision, since these individuals did not vote at their assigned subdivision.
+This could have an impact on the total number of voters on a subdivision and ward basis.
+
+There are several limitations with conducting this analysis.
+Number of subdivisions and polling location placement alone cannot solely dictate voter behaviour, as there are many other socio-economic factors which inform why an elector does or does not cast a ballot.
 

### 2023-06-03 outputs/paper/paper.qmd
Added more content to the discussion section, including measurement errors and limitations

@@
 Ward 20 (Scarborough Southwest) had the highest turnout of all Scarborough wards at 32% and had 61 subdivisions.
-Scarborough's challenges with accessible public transit are well recognized, potentially contributing to the lower voter turnout.
+Scarborough's recognized challenges with accessible public transit, paired with fewer subdivisions and polling locations potentially contribute to the lower voter turnout [@transitdesert].
 
-The creation of subdivision boundaries within wards lacks transparency and clarity.
-The City Clerk is responsible for drawing up all ward subdivisions.
-We don’t know what data and information they use to make these decisions and what types of bias may be introduced.
-Don’t know if there’s a public input process or if like creating ward boundaries, it is an individual who has little local knowledge of the very community impacted by the creation of boundaries
-
-Toronto Elections [@ElxnAccessibility] clearly outlines their process for finding and establishing accessible polling locations within subdivisions but does not elaborate on the process of establishing subdivisions themselves.
-
-There is a lack of clarity whether the people who voted at Advanced Polls or by Mail In Ballot were removed from the total number of electors by subdivision, since these individuals did not vote at their assigned subdivision.
-This could have an impact on the total number of voters on a subdivision and ward basis.
+Within the data set used for analysis, there are several potential measurement errors.
+Firstly, the raw data set listed the total number of subdivisions per ward as "NA", despite other aspects of voter turnout and subdivisions being calculated by Toronto Elections.
+Secondly, there were gaps in the number and corresponding numbering of subdivisions which were not explained by the City Clerk or Toronto Elections.
+For example, Ward 1 is listed as having 60 subdivisions, plus Advanced Vote and Mail In Voting locations, but subdivisions 4, 20, 43-46, 50, and 58 are unaccounted for, totally 51 subdivisions.
+Thirdly, there is a lack of clarity whether the people who voted at Advanced Polls or by Mail In Ballot were removed from the total number of electors by subdivision, since these individuals did not vote at their assigned subdivision.
+If voters cast their ballot early and were not removed from the total number of electors at their specific subdivision, this could impact how voter turnout by subdivision is calculated.
 
 There are several limitations with conducting this analysis.
-Number of subdivisions and polling location placement alone cannot solely dictate voter behaviour, as there are many other socio-economic factors which inform why an elector does or does not cast a ballot.
+The creation of subdivision boundaries within wards lacks transparency and clarity, while impacting the local democratic process.
+The City Clerk is responsible for drawing up all ward subdivisions, however the data and information they use to make these decisions is not public and it is challenging to account for any bias introduced during this process.
+Toronto Elections @ElxnAccessibility clearly outlines their process for finding and establishing accessible polling locations within subdivisions but does not elaborate on the process of establishing subdivisions themselves.
+
+Moreover, the number of subdivisions and polling location placement alone cannot dictate voter behaviour, as there are many other socio-economic factors which inform why an elector does or does not cast a ballot.
+The caliber of local candidates and the nature of the election, whether an incumbent is standing for re-election or the seat is open for a new councillor impacts voter turnout should be accounted for in future analysis.
 
@@
 This paper investigates the number of subdivisions within Toronto’s 25 wards in relation to voter turnout to see if the wards with more subdivisions had higher turnout during the 2022 municipal election.
-My results reveal that wards with more subdivisions did not correlate to higher voter turnout.
+My results reveal that wards with more subdivisions did not correlate to having higher voter turnout.
 Factors such as income, level of education attained, and an analysis of local races for councillor and school board trustee within a ward (whether there was an incumbent or not) should be analyzed in further studies of voter behaviour during municipal elections.

### 2023-06-03 outputs/paper/toronto_elections_paper.qmd
Updating code to graph the relationship between number of subdivisions and voter turnout

--- 2023-06-03 outputs/paper/paper.qmd
@@
 ### Turnout by Ward
-Moreover, using the cleaned voter statistics data set, I added the voter turnout by ward percentage from the bottom of each ward section to a table I created with each ward number.
+Moreover, using the cleaned voter statistics data set, I added the voter turnout by ward percentage calculated by the Toronto Elections from the bottom of each ward section to a table I created with each ward number.
 To view a sample of this data set, please see @tbl-summarized_voter_turnout.
@@
     label = "Ward"
-  ) +
-  theme_minimal()
+  ) 
 ```
@@
 Secondly, there were gaps in the number and corresponding numbering of subdivisions which were not explained by the City Clerk or Toronto Elections.
-For example, Ward 1 is listed as having 60 subdivisions, plus Advanced Vote and Mail In Voting locations, but subdivisions 4, 20, 43-46, 50, and 58 are unaccounted for, totally 51 subdivisions.
+For example, Ward 1 is listed as having 60 subdivisions, plus Advanced Vote and Mail In Voting locations, but subdivisions 4, 20, 43-46, 50, and 58 are unaccounted for, totaling 55 subdivisions.
 Thirdly, there is a lack of clarity whether the people who voted at Advanced Polls or by Mail In Ballot were removed from the total number of electors by subdivision, since these individuals did not vote at their assigned subdivision.
@@
 Moreover, the number of subdivisions and polling location placement alone cannot dictate voter behaviour, as there are many other socio-economic factors which inform why an elector does or does not cast a ballot.
-The caliber of local candidates and the nature of the election, whether an incumbent is standing for re-election or the seat is open for a new councillor impacts voter turnout should be accounted for in future analysis.
+The caliber of local candidates and the nature of the election, whether an incumbent is standing for re-election or the seat is open for a new councillor impacts voter turnout and should be accounted for in future analysis.
 

### 2023-06-03 outputs/paper/toronto_elections_paper.qmd
Updated script for graph to compare voter turnout and number of subdivisions

@@
 #Code referenced from: https://github.com/christina-wei/INF3104-1-Covid-Clinics/blob/main/outputs/paper/covid_clinics.pdf
+# & https://stackoverflow.com/questions/66652330/error-ggplot-continuous-x-aesthetic-did-you-forget-aesgroup
 finalized_voter_statistics |>
-  ggplot(aes(x = turnout, y = n, label = ward)) + 
-  geom_point() + 
+  ggplot(aes(x = factor(turnout), y = n, label = ward)) +
+  geom_point(aes(group = turnout)) + 
   geom_beeswarm() +
-  geom_smooth(method=lm, se=FALSE) + 
-  geom_text(hjust=-0.5, vjust=0.5) + 
+  geom_smooth(method="lm")+
+  geom_text(hjust=-05, vjust=0.5) + 
 labs(

### 2023-06-03 outputs/paper/toronto_elections_paper.qmd
Added abstract, fixed typos and phrasing/word choice

@@
 date-format: "long" 
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "The 2022 Toronto Municipal Election witnessed the lowest voter turnout in the city’s history since amalgamation in 1997, with voter turnout on a ward-by-ward basis ranging from 22% to 38%.
+Using data from OpenDataToronto, this paper disproves the hypothesis that having more subdivisions within a ward bolsters voter turnout.
+This paper reveals that more subdivisions within a ward does not necessarily correlate to a higher voter turnout."
 format: pdf
@@
 
-Individual turnout by ward depended on local candidates and the accessibility of subdivisions to cast a ballot.
+Individual turnout by ward depended on a number of factors, including the accessibility of polling locations within subdivisions to cast a ballot.
 The City of Toronto @ElectionDictionary defines a ward as “a geographical area represented by a member of Council.” Following Premier Doug Ford’s decision in the summer of 2018 to cut down the number of wards in the city to match federal and provincial riding boundaries, there are currently 25 wards in the City of Toronto [@BigCityElxns].
@@
 The data set used is entitled 'Elections - Voter Statistics' [@VoterStatistics].
-Data was collected and analyzed using the statistical programming software R [@r], with additional support from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], `KableExtra` [@rKableExtra], `Knitr` [@rknitr], and `here` [@rHere].
+Data was collected and analyzed using the statistical programming software R [@r], with additional support from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], `KableExtra` [@rKableExtra], `knitr` [@rknitr], `ggbeeswarm` [@rggbeeswarm], and `here` [@rHere].
 A further discussion of the data collection, cleaning, and analysis process can be found later on in this paper.
@@
 Ward 7 (Humber River-Black Creek) had the second lowest number of subdivisions at 47, following by Ward 25 (Scarborough-Rouge Park) with 48.
-@fig-voter-statistics displays the total number of subdivisions for all 25 wards.
+@fig-voter_statistics displays the total number of subdivisions for all 25 wards.
 
@@
 #| label: fig-voter_statistics
-#| fig-cap: Number of Subdivisions per Ward
+#| fig-cap: Number of Subdivisions per ward
 
@@
 #| label: fig-turnout_by_ward
-#| fig-cap: Voter Turnout by Ward
+#| fig-cap: Voter Turnout by ward
 
@@
 
-The Wards within Scarborough (Wards 20-25) consistently had some of the lowest voter turnouts and fewest number of subdivisions across all 25 wards.
+The wards within Scarborough (Wards 20-25) consistently had some of the lowest voter turnouts and fewest number of subdivisions across all 25 wards.
 Ward 23 (Scarborough North) witnessed the third lowest voter turnout across the entire city at 25% and the fewest number of subdivisions at 38.
-Ward 20 (Scarborough Southwest) had the highest turnout of all Scarborough wards at 32% and had 61 subdivisions.
+Ward 20 (Scarborough Southwest) had the highest turnout of all the Scarborough wards at 32% and had 61 subdivisions.
 Scarborough's recognized challenges with accessible public transit, paired with fewer subdivisions and polling locations potentially contribute to the lower voter turnout [@transitdesert].

### 2023-06-03 outputs/paper/toronto_elections_paper.qmd
Fixing typos and changing phrasing.

@@
 ```
-### Comparing Number of Subdivisions and Turnout
 
@@
 The creation of subdivision boundaries within wards lacks transparency and clarity, while impacting the local democratic process.
-The City Clerk is responsible for drawing up all ward subdivisions, however the data and information they use to make these decisions is not public and it is challenging to account for any bias introduced during this process.
+The City Clerk is responsible for drawing up all subdivisions, however the data and information they use to make these decisions is not public and it is challenging to account for any bias introduced during this process.
 Toronto Elections @ElxnAccessibility clearly outlines their process for finding and establishing accessible polling locations within subdivisions but does not elaborate on the process of establishing subdivisions themselves.
@@
 My results reveal that wards with more subdivisions did not correlate to having higher voter turnout.
-Factors such as income, level of education attained, and an analysis of local races for councillor and school board trustee within a ward (whether there was an incumbent or not) should be analyzed in further studies of voter behaviour during municipal elections.
+Factors such as income, level of education attained, and an analysis of local races for councillor and school board trustee within a ward should be analyzed in further studies of voter behaviour during municipal elections.
 Data from the upcoming Toronto Mayoral by-election on June 26, 2023 may reveal further insights into the placement of polling locations in relation to voter turnout.

### 2023-06-03 outputs/paper/toronto_elections_paper.qmd
Added ggrepel package to library, updated description of packages used, updated code for figure 3 to remove overlapping text, and fixed typo in discussion section.

@@
 library(ggbeeswarm)
+library(ggrepel)
 
@@
 The data set used is entitled 'Elections - Voter Statistics' [@VoterStatistics].
-Data was collected and analyzed using the statistical programming software R [@r], with additional support from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], `KableExtra` [@rKableExtra], `knitr` [@rknitr], `ggbeeswarm` [@rggbeeswarm], and `here` [@rHere].
+Data was collected and analyzed using the statistical programming software R [@r], with additional support from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], `KableExtra` [@rKableExtra], `knitr` [@rknitr], `ggbeeswarm` [@rggbeeswarm], `ggrepel` [@rggrepel], and `here` [@rHere].
 A further discussion of the data collection, cleaning, and analysis process can be found later on in this paper.
@@
 finalized_voter_statistics |>
-  ggplot(aes(x = factor(turnout), y = n, label = ward)) +
-  geom_point(aes(group = turnout)) + 
-  geom_beeswarm() +
-  geom_smooth(method="lm")+
-  geom_text(hjust=-05, vjust=0.5) + 
+  ggplot(aes(x = turnout, y = n, label = ward)) +
+  # ggplot(aes(x = factor(turnout), y = n, label = ward)) +
+  geom_point() + 
+  # geom_point(aes(group = turnout)) + 
+  # geom_beeswarm() +
+  # geom_smooth(method="lm")+
+  # geom_text(hjust=-05, vjust=0.5) +
+  ggrepel::geom_text_repel() + 
 labs(
@@
     x = "Turnout (%)",
-    y = "Subdivisions",
+    y = "Number of subdivisions",
     label = "Ward"
-  ) 
+  ) +
+  theme_classic()
 ```
@@
 
-Futhermore, Ward 7 (Humber River-Black Creek) had a 22% voter turnout and 47 subdivisions, emphasizing that having fewer subdivisions potentially contributed to having the lowest voter turnout across the city.
+Furthermore, Ward 7 (Humber River-Black Creek) had a 22% voter turnout and 47 subdivisions, emphasizing that having fewer subdivisions potentially contributed to having the lowest voter turnout across the city.
 Ward 10 (Spadina-Fork York), which had the second lowest voter turnout at 24%, accounts for the highest number of subdivisions within a ward at 94.
@@
 The City Clerk is responsible for drawing up all subdivisions, however the data and information they use to make these decisions is not public and it is challenging to account for any bias introduced during this process.
-Toronto Elections @ElxnAccessibility clearly outlines their process for finding and establishing accessible polling locations within subdivisions but does not elaborate on the process of establishing subdivisions themselves.
+Toronto Elections [@ElxnAccessibility] clearly outlines their process for finding and establishing accessible polling locations within subdivisions but does not elaborate on the process of establishing subdivisions themselves.
 
